---
title: "Mapping functional diversity"
author: "Mireia Valle"
date: "7/25/2019"
output: html_document
---

#WORKING DIRECTORY----------------------------------------------------------------------------------------------------
```{r}
getwd()
##Working on the server
setwd ("/home/valle/github/functional_traits")
##working from my laptop
#setwd("N:/github/functional_traits")
```

#LOADING Required Libraries -----------------------------------------------------------------------------------------

##Libraries for spatial analysis
```{r}
library(foreign)
library(sp)
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(maptools)
```
##Libraries for data management
```{r} 
library (tidyverse) 
library (data.table)
library (reshape)
```
##Specific libraries for getting functional traits
###FishLife
```{r}
#devtools::install_github("DanOvando/FishLife", ref= "query-fishlife")
library(FishLife)
```
###Fishbase
```{r}
#install.packages("rfishbase", 
#                 repos = c("http://packages.ropensci.org", "http://cran.rstudio.com"), 
#                 type="source")
library (rfishbase)
```

#LOADING AQUAMAPS SPECIES DATA 2018 ---------------------------------------------------------------------------------

##Loading species list 
```{r}
species <- read.csv('/home/valle/github_annex/Aquamaps/d2018/speciesoccursum_ver0816c.csv',sep=',', header = TRUE)
head (species)
summary (species)
```

#GETTING FUNCTIONAL TRAITS FROM FishLife----------------------------------------------------------------------------
##We only need to run this one time, then you go straight to the next step: *Loading Functional Traits gotten from FishLife*

##Notes on FishLife
###https://github.com/James-Thorson/FishLife
###http://onlinelibrary.wiley.com/doi/10.1002/eap.1606/full 
###* Loo = asymptotic maximum length (cm) 
###* K = individual growth rate (yr^-1) brody growth coefficient
###* Winfinity = asymptotic maximum mass (gr) 
###* tmax = maximum age (yr) 
###* tm = age at maturation (yr)
###* M = natural mortality rate (yr^-1)
###* Lm = length at maturation (cm) 
###* T = average temperature within distribution (deg C) 

##Function *Get_traits* that we are going to use was developed by Dan Ovando to query fishlife and store outputs in a sane structure, without this function the only way to extract life history parameters is by generating plots and then writing out their embedded predictions

```{r}
#?Get_traits
```
##We rename the species list data from aquamaps
```{r}
df <- species 
as.tibble (df)
```
##For a dataframe with SEPARATE 'genus' and 'species' columns (e.g., 'Thunnus','thynnus'), run the following: 
```{r}
sfg = safely(Get_traits)

species_lt2018<- df %>% 
  mutate(life_traits = map2(Genus, Species, ~(sfg(Genus=.x, Species=.y)))) %>%
  mutate(fish_life_worked = map(life_traits, 'error') %>% map_lgl(is.null)) %>%
  filter(fish_life_worked) %>%
  mutate(life_traits = map(life_traits, 'result')) %>%
  unnest() 
```
##See results
```{r}
as.tibble (species_lt2018) 
```
##See summary of the results
```{r}
summary (species_lt2018)
```
##Remove both dataframe
```{r}
rm(df)
rm (species)
```
##Save the results
```{r}
write.csv (species_lt2018, file = "species_lt2018.csv")
```

#LOADING FUNCTIONAL TRAITS GOTTEN FROM FishLife ---------------------------------------------------------------------

##Read species life traits table (species_lt2018), the table we got from FishLife
```{r}
species_lt2018 <- read.csv('species_lt2018.csv')
as.tibble (species_lt2018)
```
##We subset the table with the columns we want to keep
```{r}
species_lt2018 <- subset(species_lt2018, select=c("SPECIESID", "SpecCode", "Genus", "Species","FBname", "Kingdom", "Phylum", "Class", "Order", "Family", "Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm", "Temperature" ))
head (species_lt2018)
```
##Analysing the data
```{r}
colnames (species_lt2018)
lt_matrix <- subset(species_lt2018, select=c("Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm", "Temperature"))
lt_matrix
```
##Analysing correlations: 
```{r}
cor (lt_matrix) 
plot (lt_matrix$Loo~lt_matrix$Lm)
```

#PREPARING THE DATA TO GET SPECIES SPATIAL DISTRIBUTION -------------------------------------------------------------

##Read species life traits table (species_lt2018), the table we got from FishLife
```{r}
species_lt2018 <- read.csv('species_lt2018.csv')
as.tibble (species_lt2018)
colnames (species_lt2018)
```
##Subsetting the table to reduce the number of columns
```{r}
species_lt2018_subset <- subset(species_lt2018, select=c("SPECIESID", "Genus", "Species"))
```
##Removing the complete
```{r}
rm (species_lt2018)
```
##Keying the table to join it later
```{r}
species_lt2018_keyed <- data.table(species_lt2018_subset, key = "SPECIESID") 
```
##Removing the table that is not keyed
```{r}
rm (species_lt2018_subset)
```
#LOADING AQUAMAPS DATA ABOUT SPECIES DISTRIBUTION AND ENVIRONMENTAL DATA AND SPATIAL INFORMATION 2018----------------

##hs:table with information about the probability of occurrence of species and different environmental variables
```{r}
hs <- read.csv('/home/valle/github_annex/Aquamaps/d2018/hcaf_species_native_ver0816c.csv',sep=',', header = TRUE)
```
##See head
```{r}
head (hs)
```
##See col names in hs
```{r}
colnames (hs)
```
##Change SpeciesID by SPECIESID
```{r}
colnames (hs)<-c("SPECIESID","CsquareCode","LOICZID","probability","FAOAreaYN","BoundBoxYN","CenterLat","CenterLong")
```
##Check if the changes have been done
```{r}
colnames (hs)
head (hs)
```
##Subset hs table, we are interested in SPECIESID and CsquareCode columns only
```{r}
hs_subset <- subset(hs, select=c("SPECIESID", "CsquareCode"))
```
##Remove hs table from R environment
```{r}
rm (hs)
```
##Keying the hs table to join it later
```{r}
hs_keyed <- data.table(hs_subset,  key = "SPECIESID")
##See head
```{r}
head (hs_keyed)
```
##See which class is the table
```{r}
class(hs_keyed)
```
##Remove hs_subset table
```{r}
rm(hs_subset)
```
##We create a new table joining both tables by SpeciesID field. 
```{r}
species_hs <- hs_keyed[species_lt2018_keyed] %>%
  setkey(NULL)
```
##See head of the new table
```{r}
head (species_hs)
```
##See columns names
```{r}
colnames (species_hs)
```
##Remove the keyed tables
```{r}
rm(species_lt2018_keyed)
rm (hs_keyed)
```

#REPLACE CSquareCode with LOICZID ------------------------------------------------------
##Load xyz data (location + variables)
###xyz data frame contains information on the location of each cell and the value of diferente variables on that cell
```{r}
xyz <- read.csv('/home/valle/github_annex/Aquamaps/d2018/hcaf_v6.csv',sep='\t',header = TRUE)
```
##See head
```{r}
head (xyz)
```
##See columns names
```{r}
colnames (xyz)
```
##Create a subset with the columns we need
```{r}
xyz_subset <- subset(xyz, select=c("CsquareCode", "LOICZID"))
```
##Remove the complete table
```{r}
rm (xyz)
```
##Check columns names in the new table
```{r}
colnames (xyz_subset)
```

#We KEY both tables in order to join them later
```{r}
head (species_hs)
species_hs_keyed <- data.table(species_hs, key = "CsquareCode") 
rm(species_hs)
class (species_hs_keyed)

head (xyz_subset)
csq_loiczid <- data.table (xyz_subset, key="CsquareCode")
rm (xyz_subset)
class (csq_loiczid)
```
##We create a new table joining both tables by SpeciesID field. 
```{r}
species_loiczid <- csq_loiczid[species_hs_keyed]
head (species_loiczid)
species_loiczid <- subset(species_loiczid, select=c("LOICZID", "SPECIESID"))
head (species_loiczid)
```
##Save the results
```{r}
write.csv (species_loiczid, file = "species_loiczid.csv")
```
##Read table
```{r}
species_loiczid <- read.table ("species_loiczid.csv", head=TRUE, sep=',')
```
##See head
```{r}
head (species_loiczid)
```
##See the number of species within the table
```{r}
unique (species_loiczid$SPECIESID) %>% length()#12368
```
##See if there are any NAs
```{r}
table(is.na (species_loiczid))
```

#------------------------------------------------------------------------------------------------------------------ #
# MAPPING #
#------------------------------------------------------------------------------------------------------------------ #

## 1.- First we have to create a raster with LOICID values ----------------------------------------------------------

###To do so we generate a basemap
```{r}
basemap <- raster (ext= extent (c(-180,180, -90, 90)), res=0.5)
```

###We assign values to it
```{r}
values (basemap) <- 1:length (basemap)
basemap
```
#class       : RasterLayer 
#dimensions  : 360, 720, 259200  (nrow, ncol, ncell)
#resolution  : 0.5, 0.5  (x, y)
#extent      : -180, 180, -90, 90  (xmin, xmax, ymin, ymax)
#coord. ref. : +proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 
#data source : in memory
#names       : layer 
#values      : 1, 259200  (min, max)

###We plot it
```{r}
plot (basemap)
```

###We save it as a raster
```{r}
writeRaster(basemap, "basemap", format = "GTiff")
```

## 2.- Then we load the data.frame where we can find the variable we want to map. We will map salintity Min in this case ----------------------------------------------------------------------------------------------------------------
```{r}
xyz <- read.csv('/home/valle/github_annex/Aquamaps/d2018/hcaf_v6.csv',sep='\t',header = TRUE)
```
###We see the head and the class
```{r}
head (xyz)
class(xyz)
```
###We check the unique values of LOICZID on the data.frame
```{r}
unique (xyz$LOICZID) %>% length()
```
###We check the unique values on the map
```{r}
unique (values(basemap)) %>% length()
```

### We then create a data.frame with the two columns we want to use for mapping
```{r}
text <- xyz %>% select(LOICZID, SalinityMin)
```
###And based on that data frame we create a raster with salinity min values
```{r}
salinity <- subs (basemap, text, by= "LOICZID", which= "SalinityMin")
```
##Now we can plot it
```{r}
plot (salinity)
```

##3.- Mapping the distribution of the all species--------------------------------------------------------------------

```{r}
all_spp <-  species_loiczid %>%
  group_by (LOICZID)%>%
  summarize(n_spp = n())
  
head (all_spp)

unique (all_spp$LOICZID) %>% length()#168701
unique (all_spp$n_spp) %>% length()#3713

allspecies_map <- subs (basemap, all_spp, by= "LOICZID", which= "n_spp")
plot (allspecies_map)
```

##4.- Mapping the distribution of one species Fis-22703 -------------------------------------------------------------

```{r}
head(hs)

fis<- hs[hs$SpeciesID == "Fis-22703",]
head (fis)

fis_hs <- subset (fis, select=c("LOICZID", "probability"))
fis_map <- subs (basemap, fis_hs, by="LOICZID", which= "probability")
plot (fis_map)
```

