---
title: "Mapping functional diversity"
author: "Mireia Valle"
date: "7/25/2019"
output: html_document
---

#SETTING WORKING DIRECTORY----------------------------------------------------------------------------------------------------
```{r}
getwd()
##Working on the server
setwd ("/home/valle/github/functional_traits")
##working from my laptop
#setwd("N:/github/functional_traits")
```

#LOADING Required Libraries -----------------------------------------------------------------------------------------

##Libraries for spatial analysis
```{r}
library(foreign)
library(sp)
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(maptools)
##Libraries for data management
library (tidyverse) 
library (data.table)
library (reshape)
##Specific libraries for getting functional traits
###FishLife
#devtools::install_github("DanOvando/FishLife", ref= "query-fishlife")
library(FishLife)
###Fishbase
#install.packages("rfishbase", 
#                 repos = c("http://packages.ropensci.org", "http://cran.rstudio.com"), 
#                 type="source")
library (rfishbase)
```
#-------------------------------------------------------------------------------------------------------------------#
#LOADING AQUAMAPS SPECIES DATA 2018 AND GETTING FUNCTIONAL TRAITS FROM FishLife
##We only need to run this one time, then you go straight to the next step: #2.- LOADING FUNCTIONAL TRAITS GOTTEN FROM FishLife
#---------------------------------------------------------------------------------

##Loading species list 
```{r}
#Reading data from MAZU
#species <- read_csv ('/home/shares/ohi/git-annex/globalprep/_raw_data/aquamaps/d2018/speciesoccursum_ver0816c.csv',sep=',', header = TRUE)

#From my laptop
#species <- read.csv('/home/valle/github_annex/Aquamaps/d2018/speciesoccursum_ver0816c.csv',sep=',', header = TRUE)
#head (species)
#colnames (species)
#summary (species)
#str(species)
#unique (species$SPECIESID) %>% length()#24904
```

##Notes on FishLife
###https://github.com/James-Thorson/FishLife
###http://onlinelibrary.wiley.com/doi/10.1002/eap.1606/full 
###* Loo = asymptotic maximum length (cm) 
###* K = individual growth rate (yr^-1) brody growth coefficient
###* Winfinity = asymptotic maximum mass (gr) 
###* tmax = maximum age (yr) 
###* tm = age at maturation (yr)
###* M = natural mortality rate (yr^-1)
###* Lm = length at maturation (cm) 
###* T = average temperature within distribution (deg C) 

##Function *Get_traits* that we are going to use was developed by Dan Ovando to query fishlife and store outputs in a sane structure, without this function the only way to extract life history parameters is by generating plots and then writing out their embedded predictions

```{r}
#?Get_traits
```
##We rename the species list data from aquamaps
```{r}
#df <- species 
#as.tibble (df)
```
##For a dataframe with SEPARATE 'genus' and 'species' columns (e.g., 'Thunnus','thynnus'), run the following: 
```{r}
#sfg = safely(Get_traits)

#species_lt2018<- df %>% 
#  mutate(life_traits = map2(Genus, Species, ~(sfg(Genus=.x, Species=.y)))) %>%
#  mutate(fish_life_worked = map(life_traits, 'error') %>% map_lgl(is.null)) %>%
#  filter(fish_life_worked) %>%
#  mutate(life_traits = map(life_traits, 'result')) %>%
#  unnest() %>% 
#  mutate(sciname = paste(Genus, Species))
```
##See results
```{r}
#as.tibble (species_lt2018) 
#head (species_lt2018)
#colnames (species_lt2018)
#str(species_lt2018)
#unique (species_lt2018$SPECIESID) %>% length()#12368
#unique (species_lt2018$fish_life_worked) 
```
##We subset the table with the columns we want to keep
```{r}
#species_lt2018 <- subset(species_lt2018, select=c("SPECIESID", "SpecCode", "Genus", "Species","sciname", "FBname", #"Kingdom", "Phylum", "Class", "Order", "Family", "Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm", "Temperature" ))
#head (species_lt2018)
```
##Save the results
```{r}
#write.csv (species_lt2018, file = "species_lt2018.csv")
```
##Remove both dataframe
```{r}
#rm(df)
#rm (species)
```

#-------------------------------------------------------------------------------------------------------------------#
#2.- LOADING FUNCTIONAL TRAITS GOTTEN FROM FishLife 
#-------------------------------------------------------------------------------------------------------------------#

##Read species life traits table (species_lt2018), the table we got from FishLife
```{r}
species_lt2018 <- read.csv('species_lt2018.csv')
as.tibble (species_lt2018)
```
##Analysing the data
```{r}
colnames (species_lt2018)
lt_matrix <- subset(species_lt2018, select=c("Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm", "Temperature"))
lt_matrix
```
##Analysing correlations: 
```{r}
cor (lt_matrix) 
plot (lt_matrix$Loo~lt_matrix$Lm)
```

#-------------------------------------------------------------------------------------------------------------------#
#3.- PREPARING THE DATA TO GET SPECIES SPATIAL DISTRIBUTION 
#-------------------------------------------------------------------------------------------------------------------#

##Read species life traits table (species_lt2018), the table we got from FishLife
```{r}
#species_lt2018 <- read.csv('species_lt2018.csv')
#as.tibble (species_lt2018)
#colnames (species_lt2018)
```
##Subsetting the table to reduce the number of columns
```{r}
species_lt2018_subset <- subset(species_lt2018, select=c("SPECIESID", "sciname"))
```
##Removing the complete
```{r}
rm (species_lt2018)
```
##Keying the table to join it later
```{r}
species_lt2018_keyed <- data.table(species_lt2018_subset, key = "SPECIESID") 
```
##Removing the table that is not keyed
```{r}
rm (species_lt2018_subset)
```
#-------------------------------------------------------------------------------------------------------------------#
#4.- LOADING AQUAMAPS DATA ABOUT SPECIES DISTRIBUTION AND ENVIRONMENTAL DATA AND SPATIAL INFORMATION 2018
#-------------------------------------------------------------------------------------------------------------------#

##hs:table with information about the probability of occurrence of species and different environmental variables
```{r}
#hs <- read_csv ('/home/valle/github_annex/Aquamaps/d2018/hcaf_species_native_ver0816c.csv', col_names = TRUE, col_types = 'cc?diidd', guess_max = 100000)

#Reading data from MAZU
hs <- read_csv ('/home/shares/ohi/git-annex/globalprep/_raw_data/aquamaps/d2018/hcaf_species_native_ver0816c.csv', col_names = TRUE, col_types = 'cc?diidd', guess_max = 100000)
```
#col_types:
#c = character,
#i = integer, 
#n = number, 
#d = double, 
#l = logical, 
#f = factor, 
#D = date, 
#T = date time, 
#t = time, 
#? = guess

##See head, colnames and str
```{r}
head (hs)
colnames (hs)
str(hs)
unique (hs$SpeciesID) %>% length()#24904

```
##Change SpeciesID by SPECIESID
```{r}
colnames (hs)<-c("SPECIESID","CsquareCode","LOICZID","probability","FAOAreaYN","BoundBoxYN","CenterLat","CenterLong")
```
##Check if the changes have been done
```{r}
colnames (hs)
head (hs)
```
## See if there are NAs
```{r}
sum(is.na(hs$CsquareCode))
sum(is.na(hs$LOICZID)) #44138318
```
##Subset hs table, we are interested in SPECIESID and CsquareCode columns only
```{r}
hs_subset <- subset(hs, select=c("SPECIESID", "CsquareCode", "LOICZID","probability", "CenterLat", "CenterLong" ))
```
##Check if the changes have been done
```{r}
colnames (hs_subset)
head (hs_subset)
```

##Remove hs table from R environment
```{r}
rm (hs)
```
##Keying the hs table to join it later
```{r}
hs_keyed <- data.table(hs_subset,  key = "SPECIESID")
```
##See head
```{r}
head (hs_keyed)
```
##See which class is the table
```{r}
class(hs_keyed)
```
##Remove hs_subset table
```{r}
rm(hs_subset)
```
##We create a new table joining both tables by SpeciesID field. 
```{r}
species_hs <- hs_keyed[species_lt2018_keyed] %>%
  setkey(NULL)
```
##See head, colnames and str
```{r}
head (species_hs)
colnames (species_hs)
str(species_hs)
unique (species_hs$SPECIESID) %>% length()#12368
```
##Remove the keyed tables
```{r}
rm(species_lt2018_keyed)
rm (hs_keyed)
```
#-------------------------------------------------------------------------------------------------------------------#
#5.- REPLACE CSquareCode with LOICZID 
#-------------------------------------------------------------------------------------------------------------------#
##Load xyz data (location + variables)
###xyz data frame contains information on the location of each cell and the value of diferente variables on that cell
```{r}
#From MAZU

xyz<- read_csv ('/home/shares/ohi/git-annex/globalprep/_raw_data/aquamaps/d2018/hcaf_v6.csv',sep='\t',header = TRUE)

#From my laptop
xyz <- read.csv('/home/valle/github_annex/Aquamaps/d2018/hcaf_v6.csv',sep='\t',header = TRUE)
```
##See head, colnames and str
```{r}
head (xyz)
colnames (xyz)
str(xyz)
unique (xyz$LOICZID) %>% length()#259200
```
##Create a subset with the columns we need
```{r}
xyz_subset <- subset(xyz, select=c("CsquareCode", "LOICZID", "CenterLat", "CenterLong"))
```
##See head, colnames and str
```{r}
head (xyz_subset)
colnames (xyz_subset)
str(xyz_subset)
unique (xyz_subset$LOICZID) %>% length()#259200
```
##Remove the complete table
```{r}
rm (xyz)
```
##Check columns names in the new table
```{r}
colnames (xyz_subset)
```

#We KEY both tables in order to join them later
```{r}
head (species_hs)
species_hs_keyed <- data.table(species_hs, key = "CsquareCode") 
rm(species_hs)
class (species_hs_keyed)

head (xyz_subset)
csq_loiczid <- data.table (xyz_subset, key="CsquareCode")
rm (xyz_subset)
class (csq_loiczid)
```
##We create a new table joining both tables by SpeciesID field. 
```{r}
species_loiczid <- csq_loiczid[species_hs_keyed]
head (species_loiczid)
species_loiczid <- subset(species_loiczid, select=c("LOICZID", "SPECIESID","CenterLat", "CenterLong"))
head (species_loiczid)
```
##Save the results
```{r}
#write.csv (species_loiczid, file = "species_loiczid.csv")
```
##Read table
```{r}
species_loiczid <- read.table ("species_loiczid.csv", head=TRUE, sep=',')
```
##See head
```{r}
head (species_loiczid)
```
##See the number of species within the table
```{r}
unique (species_loiczid$SPECIESID) %>% length()#12368
```
##See if there are any NAs
```{r}
#table(is.na (species_loiczid))
```

#------------------------------------------------------------------------------------------------------------------ #
# MAPPING #
#------------------------------------------------------------------------------------------------------------------ #

## 1.- First we have to create a raster with LOICID values ----------------------------------------------------------

###To do so we generate a basemap
```{r}
basemap <- raster (ext= extent (c(-180,180, -90, 90)), res=0.5)
```

###We assign values to it
```{r}
values (basemap) <- 1:length (basemap)
basemap


loiczid_df <- rasterToPoints(basemap) %>%
  as.data.frame() %>%
  setNames(c('CenterLat', 'CenterLong', 'cell_id'))

head(loiczid_df)
as.tibble(loiczid_df)
as.tibble (species_hs_keyed)
unique (species_hs_keyed$SPECIESID) %>% length()#12368

species_df <- subset(species_hs_keyed, select=c("LOICZID", "SPECIESID","probability", "CenterLat", "CenterLong"))

joint_hs <- left_join (species_df, loiczid_df,copy = FALSE, suffix = c(".x", ".y"))
head (joint_hs)
unique (joint_hs$SPECIESID) %>% length()#12368
unique (joint_hs$LOICZID) %>% length()#168635
sum(is.na(joint_hs$LOICZID))
sum(is.na(joint_hs$SpeciesID))
sum(is.na(joint_hs$probability))
sum(is.na(joint_hs$cell_id))

write.csv (loiczid_df, file = "loiczid_df.csv")
write.csv (species_df, file = "species_df.csv")

ggplot(joint_hs, aes(x = CenterLong, y = CenterLat, fill = cell_id)) + geom_raster()

head(joint_hs)

```
#class       : RasterLayer 
#dimensions  : 360, 720, 259200  (nrow, ncol, ncell)
#resolution  : 0.5, 0.5  (x, y)
#extent      : -180, 180, -90, 90  (xmin, xmax, ymin, ymax)
#coord. ref. : +proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 
#data source : in memory
#names       : layer 
#values      : 1, 259200  (min, max)

###We plot it
```{r}
plot (basemap)
```

###We save it as a raster
```{r}
writeRaster(basemap, "basemap", format = "GTiff")
```

###Read raster
```{r}
basemap <- raster ("/home/valle/github_annex/basemap.tif")
```


## 2.- Then we load the data.frame where we can find the variable we want to map. We will map salintity Min in this case ----------------------------------------------------------------------------------------------------------------
```{r}
xyz <- read.csv('/home/valle/github_annex/Aquamaps/d2018/hcaf_v6.csv',sep='\t',header = TRUE)
```
###We see the head and the class
```{r}
head (xyz)
class(xyz)
```
###We check the unique values of LOICZID on the data.frame
```{r}
unique (xyz$LOICZID) %>% length()
```
###We check the unique values on the map
```{r}
unique (values(basemap)) %>% length()
```

### We then create a data.frame with the two columns we want to use for mapping
```{r}
text <- xyz %>% select(LOICZID, SalinityMin)
```
###And based on that data frame we create a raster with salinity min values
```{r}
salinity <- subs (basemap, text, by= "LOICZID", which= "SalinityMin")
```
##Now we can plot it
```{r}
plot (salinity)
```

##3.- Mapping the distribution of the all species--------------------------------------------------------------------

```{r}
all_spp <-  species_loiczid %>%
  group_by (LOICZID)%>%
  summarize(n_spp = n())
  
head (all_spp)

unique (all_spp$LOICZID) %>% length()#168701
unique (all_spp$n_spp) %>% length()#3713

allspecies_map <- subs (basemap, all_spp, by= "LOICZID", which= "n_spp")
plot (allspecies_map)
```

##4.- Mapping the distribution of one species Fis-22703 -------------------------------------------------------------

```{r}
head(hs)
fis1<- hs[hs$SpeciesID == "Fis-22703",]
head (fis1)

fis2<- joint_hs2[joint_hs2$SpeciesID == "Fis-22703",]
head (fis2)

fis2_hs <- subset (fis2, select=c("LOICZID", "probability"))
fis2_map <- subs (basemap, fis2_hs, by="LOICZID", which= "probability")
plot (fis2_map)
```

##PREPARING THE DATA TO CALCULATE FUNCTIONAL DIVERSITY INDEXES-------------------------------------------------------

head(species_hs)
unique (species_hs$LOICZID) %>% length()
unique (species_hs$SPECIESID) %>% length()

species_hs_subset<- subset(species_hs, select=c("SPECIESID", "LOICZID","probability"))
head (species_hs_subset)

### use spread() to turn the species ID values into column names, with the prob as the value in the cell
prueba <- read.csv('prueba.csv',sep=',', header = TRUE)
head (prueba)
prueba_matrix <- prueba %>% spread(SPECIESID, probability)
head (prueba_matrix)

species_hs_subset_id <- tibble::rowid_to_column(species_hs_subset)
head (species_hs_subset_id)
loiczid_sps_matrix <- species_hs_subset_id %>% spread(SPECIESID, probability)

###distinct() will drop any duplicated rows (keeping only one version of each distinct observation), which seems like the problem - see if spp_hs_subset has fewer rows than spp_hs

spp_hs_subset <- species_hs_subset %>%
  select(SPECIESID, LOICZID, probability) %>%
  distinct()

nrow(spp_hs_subset)
nrow (species_hs_subset)

### use spread() to turn the species ID values into column names, with the prob as the value in the cell
loiczid_sps_matrix <- spp_hs_subset %>% spread(SPECIESID, probability)

head(spp_hs_subset)

asdf <- spp_hs_subset[c(555527, 555528, 555529, 555530), ]

asdf

unique (spp_hs_subset$LOICZID) %>% length()
unique (spp_hs_subset$SPECIESID) %>% length()


colSums(is.na(species_loiczid))

total <- rbind(data frameA, data frameB) 


```{r}
species_hs_wide <- species_hs %>%
  mutate(LatLon = paste(CenterLat, CenterLong)) %>%
  select(LatLon, SpeciesID, probability) %>%
  spread(SpeciesID, probability)
```

write.csv (species_lt2018, file = "species_lt2018.csv")

```{r}
head (species_lt2018)
sp_lt2018 <- mutate(species_lt2018, sciname = paste(Genus, Species))
head (sp_lt2018)
```

jc <- hs %>% 
  filter (is.na(LOICZID))%>%
  group_by(CenterLat, CenterLong)%>%
  count ()%>%
  ggplot(aes(x=CenterLong, y=CenterLat, fill=n))
  
  
  
fishbase_traits <- read.csv('fishbase_species_select.csv',sep=',', header = TRUE)
head (fishbase_traits)
summary (fishbase_traits)

Leftjoin
?left_join
head (sp_lt2018)

spp_FT <- left_join (sp_lt2018,fishbase_traits, by = "sciname", copy = FALSE, suffix = c(".x", ".y"))

##-------------------------------------------------------------------------------------------------------------------
# CALCULATING FUNCTIONAL DIVERSITY
##-------------------------------------------------------------------------------------------------------------------

#Loading required packages
library (FD) #install.packages ("FD")

