---
title: "Fixing the messed up AquaMaps spp cell file"
author: "Casey O'Hara"
date: "8/3/2019"
output: html_document
---

```{r setup}
library(raster)
library (tidyverse) 
library (data.table)
```

### Load species list and species-cell data

```{r}
dir_am <- '/home/shares/ohi/git-annex/globalprep/_raw_data/aquamaps'

spp_cells <- data.table::fread(file.path(dir_am, 'd2018/hcaf_species_native_ver0816c.csv'))
  ### fread() way faster
spp_cells_clean <- spp_cells %>%
  setNames(tolower(names(.))) %>%
  select(am_sid = speciesid, loiczid_error = loiczid, prob = probability, lat = centerlat, long = centerlong)

head(spp_cells_clean)
```

```{r check loiczid NAs}
test_df <- spp_cells_clean %>%
  select(loiczid_error, lat, long) %>%
  distinct()

nrow(test_df)

sum(is.na(test_df$loiczid_error))
```

Length should be no larger (and certainly smaller) than 259200 (i.e. 180 * 2 * 360 * 2).  Some lat/longs have NA and numeric LOICZID.

### Create a dataframe of lat/long to LOICZID.

Create a generic raster, fill the cells horizontally from upper left to bottom right - this is equivalent to LOICZID values.  Then convert that raster to a dataframe with `lat`, `long`, and `loiczid` values.

```{r create lat long to loiczid lookup}
basemap <- raster(ext = extent (c(-180, 180, -90, 90)), 
                  res = 0.5)
values (basemap) <- 1:length (basemap)
plot(basemap)

base_df <- rasterToPoints(basemap) %>%
  as.data.frame() %>%
  setNames(c('long', 'lat', 'loiczid')) %>%
  mutate(loiczid = as.integer(loiczid))
```

Using lat and long as spatial references common to both `spp_cells` and `base_df`, join fixed LOICZID values to the `spp_cells` dataframe.  Use `data.table` keyed join!

```{r attach new loiczids to species cells}
spp_cells_dt <- spp_cells_clean %>%
  data.table(key = c('lat', 'long'))
base_dt <- base_df %>%
  data.table(key = c('lat', 'long'))

spp_cells_fixed_dt <- base_dt[spp_cells_dt]
```

Check again to see if the values are cleaned up:

```{r check loiczid NAs again}
test_df <- spp_cells_fixed_dt %>%
  select(loiczid, lat, long) %>%
  distinct()

nrow(test_df)

sum(is.na(test_df$loiczid_error))

head(spp_cells_fixed_dt)
```

That's more like it!  no NAs

### Write out the new corrected data

```{r}
spp_cells_final <- spp_cells_fixed_dt %>%
  select(am_sid, loiczid, prob) %>%
  distinct()

data.table::fwrite(spp_cells_final, 
                   file.path(dir_am, 'd2018/hcaf_species_native_ver0816c_fixed.csv'))
```

