---
title: "assemblage_matrix"
author: "Mireia Valle"
date: "8/9/2019"
output: html_document
---

```{r setup}
##Libraries for data management
library (tidyverse) 
library (data.table)
library (reshape)
```

### Load species list 
```{r}
species_lt2018 <- read.csv('species_lt2018.csv')
unique (species_lt2018$am_sid) %>% length()#12368
species_list <- species_lt2018 %>%
    select(am_sid)
unique (species_list$am_sid) %>% length()#12368
```
###Load species-cell data
```{r}
dir_am <- '/home/shares/ohi/git-annex/globalprep/_raw_data/aquamaps'

spp_cells_final <- data.table::fread(file.path(dir_am, 'd2018/hcaf_species_native_ver0816c_fixed.csv'))
  ### fread() way faster
unique (spp_cells_final$am_sid) %>% length()#24904
sum(is.na(spp_cells_final$loiczid))#0
```
###Join tables using "am_sid" common to both `spp_cells_final` and `species_lt2018` to get a table where only the data on selected marine fish species is shown. Use `data.table` keyed join.

```{r}
spp_cells_final_keyed <- spp_cells_final %>%
  data.table(key = "am_sid")
head (spp_cells_final_keyed)
unique (spp_cells_final_keyed$am_sid) %>% length()#24904
sum(is.na(spp_cells_final_keyed$loiczid))#0

species_list_keyed <- species_list %>%
  data.table(key = "am_sid")
head (species_list_keyed)
unique (species_list_keyed$am_sid) %>% length()#12368

spp_cells_selected <- spp_cells_final_keyed[species_list_keyed, nomatch = 0]

head (spp_cells_selected)
unique (spp_cells_selected$am_sid) %>% length()#12365
sum(is.na(spp_cells_selected$am_sid))
sum(is.na(spp_cells_selected$loiczid))
```

###we want to use spread() to turn the am_sid into column names, with the prob as the value in the cell

#Let's do it for the first 10 species
#select the first 10 species
first_10spp <- head(unique(spp_cells_selected$am_sid), 10)
first_10spp
#select loiczid and probability from spp_cells_selected dataframe
first_10spp_hs <- spp_cells_selected %>% filter(am_sid %in% first_10spp)
head(first_10spp_hs)
#spread the data from those first 10 species
spread_10spp <- first_10spp_hs %>%
    spread(am_sid, prob, fill = 0)
#it works!

#spread the complete dataset 
head (spp_cells_selected)
spread <- spp_cells_selected %>%
    spread(am_sid, prob, fill = 0)
    
#it doesn't work because it has duplicated key values (more than one value of probability for one species and one cell)

#use group_by () to deal with duplicated probability values per species and per cells. Using this function we group the duplicated values by mean
spp_grouped <- spp_cells_selected %>%
  group_by(am_sid, loiczid)  %>%
  summarize(prob = mean(prob))
unique (spp_grouped$am_sid) %>% length()#12365

#Now we can spread data using grouped mean values of probability for 524 cells (*2= 1048)
spread_grouped <- spp_grouped %>%
    spread(am_sid, prob, fill = 0)
head (spread_grouped)

# We can see which are those duplicated rows

duplicate <- spp_cells_selected %>%
#create a column call count where we write 1
mutate (count=1) %>%
#group by unique combinations of am_sid and loiczid
group_by (am_sid, loiczid) %>%
#summarize the data counting the values on count
summarize(sum = sum(count))%>%
#filter data selecting only that rows that have a value > 1 for that sum
filter(sum > 1)

head (duplicate)#this table contains the id of the 524 cells that had duplicated values

#we want to see how many species are involved on that duplication 
unique (duplicate$am_sid) %>% length() 
# only 2
unique (duplicate$am_sid) 
# which? "Fis-23415" "Fis-25283"

#we analyse the duplicated data for species 1
duplicate1<- duplicate[duplicate$am_sid == "Fis-23415",]
duplicate1
head (duplicate1)
#we map it
dup1_map <- subs (basemap, duplicate1, by="loiczid", which= "sum")
plot (dup1_map)
#we look for its name
duplicate1name<- species_lt2018[species_lt2018$am_sid == "Fis-23415",]
#438 duplicated probability values, species name= Carangoides equula (Temminck & Schlegel, 1844)

#we repeat the same steps for the second species
duplicate2<- duplicate[duplicate$am_sid == "Fis-25283",]
duplicate2
head (duplicate2)
#we look for its name
duplicate2name<- species_lt2018[species_lt2018$am_sid == "Fis-25283",]
#we map it
dup2_map <- subs (basemap, duplicate2, by="loiczid", which= "sum")
plot (dup2_map)

#we create a column that is call "duplicate" and is filled by "TRUE" for all rows. 
duplicate_true <- duplicate %>%
mutate (duplicate=TRUE)

#we select the columns we want to keep, we drop "count" column
duplicate_true <- subset(duplicate_true, select=c("am_sid", "loiczid","duplicate"))

#we do a left_joint to our species data set
join <- left_join (spp_cells_selected, duplicate_true)

#we check if everything went well
head (join)
#the column we have created has two unique values NA and TRUE
unique (join$duplicate) %>% length()#2
unique (join$duplicate) # NA TRUE
#We filter the joint table selecting those rows with NA values
spp_hs_final <- filter (join, is.na(duplicate))
#we check if everything went well
head (spp_hs_final)
unique (spp_hs_final$duplicate) %>% length()#1
unique (spp_hs_final$duplicate) # NA
unique (spp_hs_final$am_sid) %>% length()#

spp_hs_final <- subset(spp_hs_final, select=c("am_sid", "loiczid","prob"))
head (spp_hs_final)
as.tibble (spp_hs_final)

spread_loiczid_spp <- spp_hs_final %>%
    spread(am_sid, prob, fill = 0)

as.tibble (spread_loiczid_spp)

drops <- c("Fis-23415","Fis-25283")
spread_loiczid_spp <-spread_loiczid_spp[ , !(names(spread_loiczid_spp) %in% drops)]
as.tibble (spread_loiczid_spp)

#save the assemblage_matrix
write.csv (spread_loiczid_spp, file = "spread_loiczid_spp.csv")

