---
title: "4_fishbase_traits"
author: "Mireia Valle"
date: "8/15/2019"
output: html_document
---

```{r setup}
#installing and loading libraries

install.packages("rfishbase", 
                 repos = c("http://packages.ropensci.org", "http://cran.rstudio.com"), 
                 type="source")
library (rfishbase)
?rfishbase
library (tidyverse)
```

#read species life traits table (species_lt2018), the table we got from FishLife
```{r}
species_lt2018 <- read.csv('species_lt2018.csv')
as.tibble (species_lt2018)

species_list <- species_lt2018 %>%
select(Species=sciname)
as.tibble (species_list)
unique (species_list$Species) %>% length ()#12368
sum(is.na(species_list$Species))#0

#merge Species and Genus to create a new column call sciname
#species_list <- species_lt2018
#head (species_list)
#species_list<-unite(species_list, sciname,c(4:5), sep = " ", remove = TRUE)
#as.tibble (species_list)
#rm (species_list)
```

#START GETTING TRAIT INFORMATION FROM FISH BASE
```{r}
#-------------------------------------------------#
#species() function returns a table containing much (but not all) of the information found on the summary or homepage for a species on fishbase.org. rfishbase functions always return tidy data tables: rows are observations (e.g. a species, individual samples from a species) and columns are variables (fields).

#species###########################################################################################
#species_table<- rfishbase::docs('species')
#species_table
#write.csv(species_table, file = "species_table.csv")

#getting species data
#options(FISHBASE_API = "https://fishbase.ropensci.org/sealifebase")
options(FISHBASE_API = "https://fishbase.ropensci.org")

fishbase_species<- species (species_list$Species)
head (fishbase_species)
summary (fishbase_species)
#12368 obs of 100 variables
unique (fishbase_species$Species) %>% length ()
sum(is.na(fishbase_species$SpecCode))


#we want to add the SpecCode to our species list
species_list <- fishbase_species %>%
select(SpecCode= SpecCode, Species=Species)
as.tibble (species_list)
unique (species_list$Species) %>% length ()
sum(is.na(species_list$SpecCode))

#we want to remove the NA values from that list
spp_list <- na.omit(species_list) 
#we save the table
write.csv (spp_list, file = "spp_list.csv")

fishbase_species<- species (spp_list$Species)
head (fishbase_species)
summary (fishbase_species)
#12368 obs of 100 variables

#we select 9 of the 100 variables:

##1.- Body Shape
unique (fishbase_species$BodyShapeI) %>% length()
#7
unique (fishbase_species$BodyShapeI) 
#elongated, fusiform / normal, eel-like, short and / or deep, other, NA, other (see remarks)
sum(is.na(fishbase_species$BodyShapeI))
#1338

##2.- DemersPelag
unique (fishbase_species$DemersPelag) %>% length()
#9
unique (fishbase_species$DemersPelag) 
#1] "pelagic-oceanic" "pelagic-neritic" "demersal" "bathypelagic""bathydemersal""benthopelagic""NA""reef-associated" "pelagic"    
sum(is.na(fishbase_species$DemersPelag))
#159

#3.- DepthRangeShallow NAs: 2104
#4.- DepthRangeDeep NAs:1638
#5.- Vulnerability
sum(is.na(fishbase_species$Vulnerability))#159
#6.- Length NAs: 858 
#7.- Importance NAs: 8118
#8.- PriceCateg
#9.- PD 50 Phylogenetic diversity index (Ref. 82805):  [Uniqueness, from 0.5 = low to 2.0 = high] NAs: 6

labels (fishbase_species)
#We select those fields that we want to keep
fishbase_species_select <- fishbase_species [c("SpecCode","Species", "BodyShapeI", "DemersPelag", "DepthRangeShallow", "DepthRangeDeep", "Vulnerability", "Length", "Importance", "PriceCateg", "PD50")]
as.tibble (fishbase_species_select)
rm (fishbase_species)

#we rename the fields avoiding Up and lowe case in the labels
fishbase_species_select <- fishbase_species_select %>%
select(SpecCode=SpecCode, sciname = Species, bodyshape = BodyShapeI, demerspelag = DemersPelag, depthrangeshallow = DepthRangeShallow, depthrangedeep =DepthRangeDeep, vulnerability=Vulnerability, length =Length, importance = Importance, pricecateg = PriceCateg, pd50=PD50)

#we save the table
write.csv (fishbase_species_select, file = "fishbase_species_select.csv")

#-------------------------------------------------#
#Looking for more available data in FISH BASE

#to see documentation of tables and fields
tables <- docs()
tables
write.csv(tables, file = "tables.csv")

#diet###########################################################################################
#diet_table<- rfishbase::docs('diet')
#diet_table
#write.csv(diet_table, file = "diet_table.csv")
#getting diet data
fishbase_diet<- diet(spp_list$Species, fields=c("SpecCode", "Species", "Troph", "seTroph", "SizeMin", "SizeMax", "SizeType", "FishLength"))
head (fishbase_diet)
summary (fishbase_diet)
summary (fishbase_diet$Troph)
unique (fishbase_diet$SizeType)
as.tibble (fishbase_diet)

#THIS DATA NEED TO BE ARRANGED BEFORE BEING ADDED TO THE GLOBAL TABLE. p.e.calculate mean value by sciname

fishbase_diet_spp <- fishbase_diet %>%
  group_by(SpecCode)  %>%
  summarize(troph = mean(Troph), setrop= mean (seTroph), sizemin = mean (SizeMin), sizemax = mean (SizeMax), fishlength = mean ("FishLength", na.rm=TRUE) )

rm(fishbase_diet)
head (fishbase_diet_spp)
sum(is.na(fishbase_diet_spp$troph))
#10440
unique (fishbase_diet_spp$SpecCode) %>% length ()
#12210
sum(is.na(fishbase_diet_spp$SpecCode))#1
sum(is.na(fishbase_diet_spp$troph)) #10440
sum(is.na(fishbase_diet_spp$setrop))#10440

#we save the table
write.csv (fishbase_diet_spp, file = "fishbase_diet_spp.csv")

#ecology###########################################################################################
#ecology_table<- rfishbase::docs('ecology')
#ecology_table
#write.csv(ecology_table, file = "ecology_table.csv")
#getting ecology data
list_fields("Troph")
#we select the fields we want to keep
fishbase_ecol<- ecology (spp_list$Species, fields=c("SpecCode", "Epipelagic", "Mesopelagic", "Bathypelagic", "Abyssopelagic", "Hadopelagic", "FeedingType", "DietTroph", "DietSeTroph", "FoodTroph", "FoodSeTroph","FoodRemark", "AddRems", "Benthic", "Sessile", "Mobile", "Demersal", "Endofauna", "Pelagic"))
head (fishbase_ecol)
summary (fishbase_ecol)
#12378 obs of 19 variables
#"DietTroph" NAs: 10905
#"DietSeTroph" NAs 11153
#"FoodTroph" NAs 7504
#"FoodSeTroph" NAs 7504
#"FoodRemark"
#"AddRems"

unique (fishbase_ecol$SpecCode) %>% length ()
#12210
sum(is.na(fishbase_ecol$SpecCode))
#159
sum(is.na(fishbase_ecol$DietTroph)) #10905
sum(is.na(fishbase_ecol$DietSeTroph))#11153
sum(is.na(fishbase_ecol$FoodTroph)) #7504
sum(is.na(fishbase_ecol$FoodSeTroph))#7504
sum(is.na(fishbase_ecol$FeedingType))#7640

fishbase_ecol_spp <- fishbase_ecol %>%
  group_by(SpecCode)  %>%
  summarize(Epipelagic = first (Epipelagic), Mesopelagic= first (Mesopelagic), Bathypelagic= first (Bathypelagic), Abyssopelagic= first (Abyssopelagic), Hadopelagic = first(Hadopelagic), FeedingType = first (FeedingType), DietTroph = first (DietTroph), DietSeTroph =first (DietSeTroph), FoodTroph = first (FoodTroph), FoodSeTroph = first (FoodSeTroph), FoodRemark = first (FoodRemark), AddRems = first (AddRems), Benthic = first (Benthic), Sessile = first (Sessile), Mobile = first (Mobile),  Demersal = first (Demersal), Endofauna = first (Endofauna), Pelagic = first (Pelagic))
    

#ecol_na <-fishbase_ecol[fishbase_ecol$SpecCode == "NA",]
#head (ecol_na)

#we save the table
write.csv (fishbase_ecol_spp, file = "fishbase_ecol.csv")

#fecundity###########################################################################################
#fecun_table<- rfishbase::docs('fecundity')
#fecun_table
#write.csv(fecun_table, file = "fecun_table.csv")
#getting fecundity data
#fecundity ("Eleginus nawaga")
fishbase_fecun <- fecundity (spp_list$Species, fields=c("SpecCode", "FecundityType", "SourceRef", "FecundityMin", "FecundityMax"))
head (fishbase_fecun)
summary (fishbase_fecun)
unique (fishbase_fecun$SpecCode) %>% length ()#12210

fishbase_fecun_spp <- fishbase_fecun %>%
  group_by(SpecCode)  %>%
  summarize(fecuntype= first (FecundityType), fecun_min = mean(FecundityMin), fecun_max= mean (FecundityMax) )

head (fishbase_fecun_spp)
unique (fishbase_fecun_spp$SpecCode) %>% length ()
#12210
sum(is.na(fishbase_fecun_spp$SpecCode))
#1
sum(is.na(fishbase_fecun_spp$fecuntype)) #12161
sum(is.na(fishbase_fecun_spp$fecun_min)) #11603
sum(is.na(fishbase_fecun_spp$fecun_max)) #11535

rm (fishbase_fecun)
#we save the table
write.csv (fishbase_fecun_spp, file = "fishbase_fecun_spp.csv")

#length###########################################################################################
#e<- length_length ("Pomadasys stridens")
#head (e)
#getting length data
#fishbase_length <- length_length (species_list$sciname, fields=c("SpecCode","LengthMin", "LengthMax"))
#head (fishbase_length)
#summary(fishbase_length)
#calculate mean value by sciname
#16974 obs of 2 variables 
#LengthMin 
#LengthMax

#maturity###########################################################################################
#matu_table<- rfishbase::docs('maturity')
#matu_table
#write.csv(matu_table, file = "matu_table.csv")
#getting maturity data
#fishbase_matu <- maturity (species_list$sciname, fields=c("SpecCode", "tm", "Lm"))
#Error in curl::curl_fetch_memory(url, handle = handle) : 
#Timeout was reached: Connection timed out after 10000 milliseconds
#head (fishbase_matu)
#summary (fishbase_matu)

#morphology###########################################################################################
#morpho_table<- rfishbase::docs('morphology')
#morpho_table
#write.csv(morpho_table, file = "morpho_table.csv")
#getting morphology data
fishbase_morpho <- morphology (spp_list$Species)
head (fishbase_morpho)
summary (fishbase_morpho)

unique (fishbase_morpho$SpecCode) %>% length ()
#12210
sum(is.na(fishbase_morpho$SpecCode))
#159

labels(fishbase_morpho)

fishbase_morpho_select  <- fishbase_morpho [c("SpecCode","OperculumPresent", "BodyShapeI", "SexualAttributes", "SexMorphology", "TypeofEyes", "TypeofMouth", "PosofMouth", "CaudalFinI", "CShape", "PelvicsAttributes")]

sum(is.na(fishbase_morpho_select$OperculumPresent))
#3037
sum(is.na(fishbase_morpho_select$BodyShapeI))
#3796
sum(is.na(fishbase_morpho_select$SexualAttributes))
#11501
sum(is.na(fishbase_morpho_select$SexMorphology))
#11524
sum(is.na(fishbase_morpho_select$TypeofEyes))
#9952
sum(is.na(fishbase_morpho_select$TypeofMouth))
#9822
sum(is.na(fishbase_morpho_select$PosofMouth))
#9708
sum(is.na(fishbase_morpho_select$CaudalFinI))
#10540
sum(is.na(fishbase_morpho_select$CShape))
#8027
sum(is.na(fishbase_morpho_select$PelvicsAttributes))
#8529

fishbase_morpho_select_spp <- fishbase_morpho_select %>%
  group_by(SpecCode)  %>%
  summarize(OperculumPresent = first (OperculumPresent), BodyShapeI= first (BodyShapeI), SexualAttributes= first (SexualAttributes), SexMorphology= first (SexMorphology), TypeofEyes = first(TypeofEyes), TypeofMouth = first (TypeofMouth), PosofMouth = first (PosofMouth), CaudalFinI =first (CaudalFinI), CShape = first (CShape), PelvicsAttributes = first (PelvicsAttributes))

rm (fishbase_morpho)
#we save the table
write.csv (fishbase_morpho_select_spp, file = "fishbase_morpho_select.csv")

#popchar####################################################################
#popchar_table<- rfishbase::docs('popchar')
#popchar_table
#write.csv(popchar_table, file = "popchar_table.csv")
#getting popchar data
#fishbase_pop <- popchar (species_list$sciname, fields= c("SpecCode", "Wmax", "Lmax", "tmax", "SameWL", "SameLt"))
#head (fishbase_pop)
#summary (fishbase_pop) 

#popgrowth####################################################################
#popgrowth_table<- rfishbase::docs('popgrowth')
#popgrowth_table
#write.csv(popgrowth_table, file = "popgrowth_table.csv")
#getting popgrowth data
#fishbase_popgrowth <- popgrowth (species_list$sciname)
#head (fishbase_popgrowth)
#summary (fishbase_popgrowth) 

#poplf####################################################################
#poplf_table<- rfishbase::docs('poplf')
#poplf_table
#write.csv(poplf_table, file = "poplf_table.csv")
#getting poplf data
#fishbase_poplf <- poplf (species_list$sciname)
#head (fishbase_poplf)
#summary (fishbase_poplf) 

#popll####################################################################
#popll_table<- rfishbase::docs('popll')
#popll_table
#write.csv(popll_table, file = "popll_table.csv")
#getting poplf data
#fishbase_popll <- popll (species_list$sciname, fields= c("SpecCode", "Length1", "Length2", "a", "b", "LengthMin", "LengthMax"))
#head (fishbase_popll)
#summary (fishbase_popll) 

#poplw####################################################################
#poplw_table<- rfishbase::docs('poplw')
#poplw_table
#write.csv(poplw_table, file = "poplw_table.csv")

#popqb####################################################################
#popqb_table<- rfishbase::docs('popqb')
#popqb_table
#write.csv(popqb_table, file = "popqb_table.csv")
#getting popqb data
#fishbase_popqb<- popqb (species_lt2018$sciname, fields= c("SpecCode", "PopQB", "MaintQB", "Winf", "FoodType", "K", "Salinity", "t0", "Temperature", "Mortality"))
#head (fishbase_popqb)
#summary (fishbase_popqb) 

#Stocks###########################################################################################
#stocks_table<- rfishbase::docs('stocks')
#stocks_table
#write.csv(stocks_table, file = "stocks_table.csv")
#getting stocks data
fishbase_stocks <- stocks(spp_list$Species, fields= c("SpecCode", "IUCN_Code", "Protected", "TempMin", "TempMax", "TempRef", "EnvTemp", "Resilience", "ResilienceRemark"))
head (fishbase_stocks)
summary (fishbase_stocks)

#THIS DATA NEED TO BE ARRANGED BEFORE BEING ADDED TO THE GLOBAL TABLE. p.e.calculate mean value by sciname
fishbase_stocks_spp <- fishbase_stocks %>%
  group_by(SpecCode) %>%
  summarize(IUCN_Code = first(IUCN_Code), Protected = first(Protected), EnvTemp= first(EnvTemp), Resilience= first(Resilience), ResilienceRemark= first(ResilienceRemark))
  
head (fishbase_stocks_spp)
sum(is.na(fishbase_stocks_spp$IUCN_Code)) #1
sum(is.na(fishbase_stocks_spp$Resilience)) #835
sum(is.na(fishbase_stocks_spp$EnvTemp)) #1

unique (fishbase_stocks_spp$SpecCode) %>% length ()
#12210
sum(is.na(fishbase_diet_spp$SpecCode))
#1
unique (fishbase_stocks_spp$EnvTemp) 

rm (fishbase_stocks)
#we save the table
write.csv (fishbase_stocks_spp, file = "fishbase_stocks_spp.csv")

#Swimming###########################################################################################
#swim_table<- rfishbase::docs('swimming')
#swim_table
#write.csv(swim_table, file = "swim_table.csv")
#getting swimming data
fishbase_swim <- swimming (spp_list$Species, fields=c("SpecCode", "AdultType", "AdultMode"))
head (fishbase_swim)
summary (fishbase_swim)
sum(is.na(fishbase_swim$AdultType))
unique (fishbase_swim$SpecCode) %>%length ()#12210
sum(is.na(fishbase_swim$SpecCode))#159
sum(is.na(fishbase_swim$AdultType))#10422
sum(is.na(fishbase_swim$AdultMode)) #10424

#we save the table
write.csv (fishbase_swim, file = "fishbase_swim.csv")

```

#Joining all tables we got from FISHBASE
```{r}
spp_list <- read.csv ("spp_list.csv")
fishbase_species_select <- read.csv ('fishbase_species_select.csv')
fishbase_diet_spp <- read.csv ('fishbase_diet_spp.csv')
fishbase_ecol <- read.csv('fishbase_ecol.csv')
fishbase_fecun_spp <- read.csv('fishbase_fecun_spp.csv')
fishbase_morpho_select <- read.csv('fishbase_morpho_select.csv')
fishbase_stocks_spp <- read.csv('fishbase_stocks_spp.csv')
fishbase_swim <- read.csv('fishbase_swim.csv')

#view tables
as.tibble (spp_list)
as.tibble(fishbase_species_select)
as.tibble(fishbase_diet_spp)
as.tibble(fishbase_ecol)
as.tibble(fishbase_fecun_spp)
as.tibble(fishbase_morpho_select)
as.tibble(fishbase_stocks_spp)
as.tibble(fishbase_swim)

fishbase_traits <- spp_list %>%
  left_join (.,fishbase_species_select, by= "SpecCode") %>%
  left_join (.,fishbase_diet_spp, by= "SpecCode") %>%
  left_join (.,fishbase_ecol, by= "SpecCode") %>%
  left_join (.,fishbase_fecun_spp, by= "SpecCode") %>%  
  left_join (.,fishbase_morpho_select, by= "SpecCode") %>%    
  left_join (.,fishbase_stocks_spp, by= "SpecCode") %>% 
  left_join (.,fishbase_swim, by= "SpecCode")
  
fishbase_traits <- fishbase_traits %>%
  select(SpecCode = SpecCode, Species = Species, sciname = sciname, bodyshape = bodyshape, demerspelag = demerspelag, depthrangeshallow = depthrangeshallow, depthrangedeep = depthrangedeep, vulnerability = vulnerability, length= length, importance= importance, pricecateg= pricecateg, pd50= pd50, troph= troph, setrop= setrop, sizemin= sizemin, sizemax= sizemax, fishlength= fishlength, Epipelagic= Epipelagic, Mesopelagic= Mesopelagic, Bathypelagic= Bathypelagic, Abyssopelagic= Abyssopelagic, Hadopelagic= Hadopelagic, FeedingType= FeedingType, DietTroph= DietTroph, DietSeTroph= DietSeTroph, FoodTroph= FoodTroph,FoodSeTroph= FoodSeTroph,FoodRemark=FoodRemark, AddRems= AddRems,Benthic= Benthic, Sessile=Sessile,  Mobile=  Mobile, Demersal= Demersal, Endofauna= Endofauna, Pelagic= Pelagic,  fecuntype= fecuntype, fecun_min= fecun_min, fecun_max= fecun_max, OperculumPresent= OperculumPresent, BodyShapeI= BodyShapeI, SexualAttributes= SexualAttributes, SexMorphology= SexMorphology, TypeofEyes=TypeofEyes,PosofMouth= PosofMouth,TypeofMouth= TypeofMouth, CaudalFinI=CaudalFinI,  CShape=CShape, PelvicsAttributes= PelvicsAttributes,  IUCN_Code=IUCN_Code, Protected= Protected, EnvTemp=EnvTemp, Resilience= Resilience, ResilienceRemark= ResilienceRemark, AdultType= AdultType, AdultMode= AdultMode)

#we save the table
write.csv (fishbase_traits, file = "fishbase_traits.csv")
```

#Creating the traits matrix joining the life traits data frame with the fishbase dataset

setwd("~/github_annex")
species_lt2018 <- read.csv('species_lt2018.csv')
unique (species_lt2018$am_sid) %>% length ()
sum(is.na(species_lt2018$am_sid))

#we want to remove the NA values from that list
spp_list <- na.omit(species_list) 
#we save the table
write.csv (spp_list, file = "spp_list.csv")

fishbase_traits <- read.csv ("fishbase_traits.csv")

