---
title: "traits_matrix"
author: "Mireia Valle"
date: "8/19/2019"
output: html_document
---

```{r setup}
#installing and loading libraries
library (tidyverse)
```

#Creating the traits matrix joining the life traits data frame with the fishbase dataset
```{r setup}
setwd("~/github_annex")
species_lt2018 <- read.csv('species_lt2018.csv')
unique (species_lt2018$am_sid) %>% length () #12368
sum(is.na(species_lt2018$am_sid))

fishbase_traits <- read.csv ("fishbase_traits.csv")
as.tibble (fishbase_traits)
unique (fishbase_traits$sciname) %>% length () #12209
sum(is.na(fishbase_traits$sciname)

traits_matrix <- left_join (fishbase_traits, species_lt2018, by= "sciname")
as.tibble (traits_matrix)
unique (traits_matrix$sciname) %>% length () #12209
sum(is.na(traits_matrix$sciname))

summary (traits_matrix)

labels (traits_matrix)

traits_matrix <- traits_matrix %>%
  select(SpecCode = SpecCode, Species = Species, sciname = sciname, bodyshape = bodyshape, demerspelag = demerspelag, depthrangeshallow = depthrangeshallow, depthrangedeep = depthrangedeep, vulnerability = vulnerability, length= length, importance= importance, pricecateg= pricecateg, pd50= pd50, troph= troph, setrop= setrop, sizemin= sizemin, sizemax= sizemax, fishlength= fishlength, Epipelagic= Epipelagic, Mesopelagic= Mesopelagic, Bathypelagic= Bathypelagic, Abyssopelagic= Abyssopelagic, Hadopelagic= Hadopelagic, FeedingType= FeedingType, DietTroph= DietTroph, DietSeTroph= DietSeTroph, FoodTroph= FoodTroph,FoodSeTroph= FoodSeTroph,FoodRemark=FoodRemark, AddRems= AddRems,Benthic= Benthic, Sessile=Sessile,  Mobile=  Mobile, Demersal= Demersal, Endofauna= Endofauna, Pelagic= Pelagic,  fecuntype= fecuntype, fecun_min= fecun_min, fecun_max= fecun_max, OperculumPresent= OperculumPresent, BodyShapeI= BodyShapeI, SexualAttributes= SexualAttributes, SexMorphology= SexMorphology, TypeofEyes=TypeofEyes,PosofMouth= PosofMouth,TypeofMouth= TypeofMouth, CaudalFinI=CaudalFinI,  CShape=CShape, PelvicsAttributes= PelvicsAttributes,  IUCN_Code=IUCN_Code, Protected= Protected, EnvTemp=EnvTemp, Resilience= Resilience, ResilienceRemark= ResilienceRemark, AdultType= AdultType, AdultMode= AdultMode, am_sid = am_sid, genus = genus, species = species, kingdom = kingdom, phylum = phylum, class = class, order = order, family = family, Loo = Loo, K = K, Winfinity = Winfinity, tmax = tmax, tm = tm, M= M, Lm = Lm, temperature= Temperature)

#find the species that we removed from the assemblage matrix
fis1<- traits_matrix[traits_matrix$am_sid == "Fis-23415",]
head (fis1)

fis2<- traits_matrix[traits_matrix$am_sid == "Fis-25283",]
head (fis2)

#remove those species from the matrix
traits_matrix_ok <- traits_matrix[-c(1937,3173), ]

#we save the table
write.csv (traits_matrix_ok, file = "traits_matrix.csv")

#load traits matrix
traits_matrix <- read.csv ("traits_matrix.csv")

as.tibble (traits_matrix)

summary (traits_matrix)
```
SpecCode = SpecCode
Species = Species
sciname = sciname
bodyshape = bodyshape
demerspelag = demerspelag
depthrangeshallow = depthrangeshallow
depthrangedeep = depthrangedeep
vulnerability = vulnerability
length= length
importance= importance
pricecateg <fct>
pd50 <dbl>
troph <dbl>
setrop <dbl>
sizemin <dbl>
sizemax <dbl>
fishlength <lgl>
Epipelagic <int>
Mesopelagic <int>
Bathypelagic <int>
Abyssopelagic <int>
Hadopelagic <int>
FeedingType <fct>
DietTroph <dbl>
DietSeTroph <dbl>
FoodTroph <dbl>
FoodSeTroph <dbl>
FoodRemark <fct>
AddRems <fct>
Benthic <int>
Sessile <int>
Mobile <int>
Demersal <int>
Endofauna <int>
Pelagic <int>
fecuntype <fct>
fecun_min <dbl>
fecun_max <dbl>
OperculumPresent <int>
BodyShapeI <fct>
SexualAttributes <fct>
SexMorphology <fct>
TypeofEyes <fct>
PosofMouth <fct>
TypeofMouth <fct>
CaudalFinI <fct>
CShape <fct>
PelvicsAttributes <fct>
IUCN_Code <fct>
Protected <int>
EnvTemp <fct>
Resilience <fct>
ResilienceRemark <fct>
AdultType <fct>
AdultMode <fct>
am_sid <fct>
genus <fct>
species <fct>
kingdom <fct>
phylum <fct>
class <fct>
order <fct>
family <fct>
Loo <dbl>
K <dbl>
Winfinity <dbl>
tmax <dbl>
tm <dbl>
M <dbl>
Lm <dbl>
temperature <dbl>


#selected traits
```{r setup}
traits_selected <- subset(traits_matrix, select=c("SpecCode", "sciname", "am_sid", "genus", "species", "kingdom", "phylum", "class", "order", "family", "Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm", "temperature",  "bodyshape", "demerspelag", "depthrangedeep", "vulnerability", "importance", "pricecateg", "pd50", "troph", "setrop", "FeedingType", "DietTroph", "DietSeTroph", "FoodTroph", "FoodSeTroph", "fecuntype", "fecun_min", "fecun_max", "OperculumPresent", "BodyShapeI", "SexualAttributes", "SexMorphology", "TypeofEyes", "PosofMouth", "TypeofMouth", "CaudalFinI", "CShape", "PelvicsAttributes", "IUCN_Code", "Protected", "EnvTemp", "Resilience", "AdultType", "AdultMode"))

#we save the table
write.csv (traits_selected, file = "traits_selected.csv")
```