---
title: "7.2_Bay of Bizcay"
author: "Mireia Valle"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#BUILDING A DATA FRAME WITH LOICZID VALUE EQUAL TO THE EXTENT WE ARE INTERESTED IN 
##We first create the global map
```{r, include = FALSE}

basemap <- raster (ext= extent (c(-180,180, -90, 90)), res=0.5)
values (basemap) <- 1:length (basemap)
basemap
plot (basemap)
```
##We then create a data frame with long, lat and loiczid columns
```{r, include = FALSE}

base_df <- rasterToPoints(basemap) %>%
  as.data.frame() %>%
  setNames(c('long', 'lat', 'loiczid')) %>%
  mutate(loiczid = as.integer(loiczid))
```
##We filter the data frame to the extent we are interested in: (ext= extent (c(-3,0, 43, 44))
```{r setup}
zoom_base_df1 <- filter(base_df, between(base_df$long, -3, 1))
zoom_base_df <- filter(zoom_base_df1, between(zoom_base_df1$lat, 43, 44))
```
##We remove the first filtered dataframe
```{r setup}
rm(zoom_base_df1)
```
##We look at our new zoomed data frame
```{r setup}
head (zoom_base_df)
summary (zoom_base_df$loiczid)
unique (zoom_base_df$loiczid) %>% length ()#16
summary (zoom_base_df$lat)
summary (zoom_base_df$long)
```
##We create a vector with loiczid values for the extent we are interested in
```{r setup}
zoom_loiczid<- zoom_base_df$loiczid
```
##We filter "species_hs" dataframe in order to get whole information about the different species that are located in the are we have selected
```{r setup}
#species_hs <- run 7_Mapping traits.Rmd from line 38 to line 76
head (species_hs)
```
##We filter the spcies_hs dataframe by the loiczid that we want to analyse
```{r setup}
zoom_species_hs <- filter (species_hs, species_hs$loiczid %in% zoom_loiczid)
unique (zoom_species_hs$loiczid) %>% length() #7
unique (zoom_species_hs$am_sid) %>% length()#784

occurrence <- filter(zoom_species_hs, prob ==1)
unique (occurrence$am_sid) %>% length()#342

head (occurrence)
colnames (occurrence)
str (occurrence)

zoom_species_hs <- occurrence
rm (occurrence)
```
##Filter is keeping all levels so, we create a new dataframe to avoid that
```{r setup, include=FALSE}
zoom_species_hs_ok<-as.data.frame( matrix(NA, nrow(zoom_species_hs), ncol(zoom_species_hs), 
                  dimnames=list( row.names(zoom_species_hs), names(zoom_species_hs) ) ) ) 
#am_sid
zoom_species_hs_ok [,"am_sid"]<-as.character(zoom_species_hs [,"am_sid"])
zoom_species_hs_ok[,"am_sid"]<-factor(zoom_species_hs_ok[,"am_sid"], ordered = TRUE)
#loiczid
zoom_species_hs_ok [,"loiczid"]<-as.integer(zoom_species_hs [,"loiczid"])
#prob
zoom_species_hs_ok[,"prob"]<-as.numeric(zoom_species_hs[,"prob"] )
#class
zoom_species_hs_ok [,"class"]<-as.character(zoom_species_hs [,"class"])
zoom_species_hs_ok[,"class"]<-factor(zoom_species_hs_ok[,"class"], ordered = TRUE)
#order
zoom_species_hs_ok [,"order"]<-as.character(zoom_species_hs [,"order"])
zoom_species_hs_ok[,"order"]<-factor(zoom_species_hs_ok[,"order"], ordered = TRUE)
#family
zoom_species_hs_ok [,"family"]<-as.character(zoom_species_hs [,"family"])
zoom_species_hs_ok[,"family"]<-factor(zoom_species_hs_ok[,"family"], ordered = TRUE)
#Loo
zoom_species_hs_ok[,"Loo"]<-as.numeric( zoom_species_hs[,"Loo"] )
#K
zoom_species_hs_ok[,"K"]<-as.numeric( zoom_species_hs[,"K"] )
#Winfinity
zoom_species_hs_ok[,"Winfinity"]<-as.numeric( zoom_species_hs[,"Winfinity"] )
#tmax
zoom_species_hs_ok[,"tmax"]<-as.numeric( zoom_species_hs[,"tmax"] )
#tm
zoom_species_hs_ok[,"tm"]<-as.numeric( zoom_species_hs[,"tm"] )
#M
zoom_species_hs_ok[,"M"]<-as.numeric( zoom_species_hs[,"M"] )
#Lm
zoom_species_hs_ok[,"Lm"]<-as.numeric( zoom_species_hs[,"Lm"] )
#depthrangedeep
zoom_species_hs_ok[,"depthrangedeep"]<-as.numeric( zoom_species_hs[,"depthrangedeep"] )
#demerspelag  
zoom_species_hs_ok [,"demerspelag"]<-as.character(zoom_species_hs [,"demerspelag"])
zoom_species_hs_ok[,"demerspelag"]<-factor(zoom_species_hs_ok[,"demerspelag"], ordered = TRUE)
#EnvTemp       
zoom_species_hs_ok [,"EnvTemp"]<-as.character(zoom_species_hs [,"EnvTemp"])
zoom_species_hs_ok[,"EnvTemp"]<-factor(zoom_species_hs_ok[,"EnvTemp"], ordered = TRUE)
#FeedingType   
zoom_species_hs_ok [,"FeedingType"]<-as.character(zoom_species_hs [,"FeedingType"])
zoom_species_hs_ok[,"FeedingType"]<-factor(zoom_species_hs_ok[,"FeedingType"], ordered = TRUE)
#bodyshape     
zoom_species_hs_ok [,"bodyshape"]<-as.character(zoom_species_hs [,"bodyshape"])
#we rename the levels into new categories names
table (zoom_species_hs_ok$bodyshape)

#zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="eel-like"),"bodyshape"]<-"eel_like"
#zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="elongated"),"bodyshape"]<-"elongated"
#zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="fusiform / normal"),"bodyshape"]<-"fusiform_normal"
#zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="short and / or deep"),"bodyshape"]<-"short_andor_deep"
#zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="other"),"bodyshape"]<-"other"

#zoom_species_hs_ok[,"bodyshape"]<-factor(zoom_species_hs_ok[,"bodyshape"], levels=c("eel_like", "elongated", "fusiform_normal", "short_andor_deep", "other" ), ordered = TRUE )

#we rename into numbers
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="eel-like"),"bodyshape"]<-"1"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="elongated"),"bodyshape"]<-"2"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="fusiform / normal"),"bodyshape"]<-"3"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="short and / or deep"),"bodyshape"]<-"4"

zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="other"),"bodyshape"]<-"5"

zoom_species_hs_ok[,"bodyshape"]<-factor(zoom_species_hs_ok[,"bodyshape"], levels=c("1", "2", "3", "4", "5" ), ordered = TRUE )
table (zoom_species_hs_ok$bodyshape)

#IUCN_Code     
zoom_species_hs_ok [,"IUCN_Code"]<-as.character(zoom_species_hs [,"IUCN_Code"])
zoom_species_hs_ok[,"IUCN_Code"]<-factor(zoom_species_hs_ok[,"IUCN_Code"], ordered = TRUE)
#Resilience    
zoom_species_hs_ok [,"Resilience"]<-as.character(zoom_species_hs [,"Resilience"])
zoom_species_hs_ok[,"Resilience"]<-factor(zoom_species_hs_ok[,"Resilience"], ordered = TRUE)
#pd50          
zoom_species_hs_ok[,"pd50"]<-as.numeric( zoom_species_hs[,"pd50"] ) 
#vulnerability 
zoom_species_hs_ok[,"vulnerability"]<-as.numeric( zoom_species_hs[,"vulnerability"] ) 
#importance    
zoom_species_hs_ok [,"importance"]<-as.character(zoom_species_hs [,"importance"])
zoom_species_hs_ok[,"importance"]<-factor(zoom_species_hs_ok[,"importance"], ordered = TRUE) 
#pricecateg    
zoom_species_hs_ok [,"pricecateg"]<-as.character(zoom_species_hs [,"pricecateg"])
zoom_species_hs_ok[,"pricecateg"]<-factor(zoom_species_hs_ok[,"pricecateg"], ordered = TRUE)

str (zoom_species_hs_ok)

```
##We remove the original data filtered dataset
```{r setup}
rm (zoom_species_hs)
```
##We create a subset of the zoom_species_hs_ok in order to create the assemblage matrix (loiczid/spps probability)
```{r setup}
zoom_species_hs_subset <- subset(zoom_species_hs_ok , select=c("am_sid", "loiczid","prob"))

head (zoom_species_hs_subset)
as.tibble (zoom_species_hs_subset)

unique (zoom_species_hs_subset$am_sid) %>% length()#342

#we spread the data
zoom_spread_loiczid_spp <- zoom_species_hs_subset %>%
    spread(am_sid, prob, fill = 0)

as.tibble (zoom_spread_loiczid_spp)
#we save the data
#write.csv (zoom_spread_loiczid_spp, file = "zoom_assemblage_eh.csv")
```
#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF ALL SPECIES IN THE SELECTED EXTEND (1214 spp)
#------------------------------------------------------------------------------------- #
```{r setup}
zoom_spp <-  zoom_species_hs_ok %>%
  group_by (loiczid)%>%
  summarize(n_spp = n())
  
head (zoom_spp)
```
##We create the zoom map
```{r setup}
zoom_basemap <- raster (ext= extent (c(-3,1, 43, 44)), res=0.5)
values (zoom_basemap) <- zoom_loiczid
plot (zoom_basemap)

zoom_spp_map <- subs (zoom_basemap, zoom_spp, by= "loiczid", which= "n_spp")

plot (zoom_spp_map)
```
#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF demerspelag TRAIT RICHNESS
#------------------------------------------------------------------------------------- #
```{r setup}
zoom_demerspelag <- zoom_species_hs_ok %>%
  group_by (loiczid)%>%
  summarize(n_distinct(demerspelag))

head (zoom_demerspelag)

zoom_demerspelag_map <- subs (zoom_basemap, zoom_demerspelag, by= "loiczid", which= "n_distinct(demerspelag)")
plot (zoom_demerspelag_map)
breakpoints <- c(0,1,2,3)
colors <- c("yellow","green","orange","red")
plot(zoom_demerspelag_map,breaks=breakpoints,col=colors)
```
#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF EnvTemp TRAIT RICHNESS
#------------------------------------------------------------------------------------- #
```{r setup}
table (zoom_species_hs_ok$EnvTemp)

zoom_EnvTemp <- zoom_species_hs_ok %>%
  group_by (loiczid)%>%
  summarize(n_distinct(EnvTemp))

head (zoom_EnvTemp)

zoom_EnvTemp_map <- subs (zoom_basemap, zoom_EnvTemp, by= "loiczid", which= "n_distinct(EnvTemp)")
breakpoints <- c(0,1,2,3,4,5,6)
colors <- c("yellow","green","orange","red", "blue", "grey")
plot(zoom_EnvTemp_map,breaks=breakpoints,col=colors)
```
#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF Loo TRAIT RICHNESS
#------------------------------------------------------------------------------------- #
```{r setup}
#Summarize Loo
summary (zoom_species_hs_ok$Loo)

#Create 4 different group intervals
zoom_species_hs_ok$Loo_groups<-cut(zoom_species_hs_ok$Loo, c(0,25,50,100,1500), labels=c("1", "2", "3", "4"))

table (zoom_species_hs_ok$Loo_groups)

head (zoom_species_hs_ok)

summary (zoom_species_hs_ok$Loo_groups)

zoom_Loo <- zoom_species_hs_ok %>%
  group_by (loiczid)%>%
  summarize(n_distinct(Loo_groups))

zoom_Loo_map <- subs (zoom_basemap, zoom_Loo, by= "loiczid", which= "n_distinct(Loo_groups)")
breakpoints <- c(0,1,2,3,4)
colors <- c("yellow","green","orange","red")
plot(zoom_Loo_map,breaks=breakpoints,col=colors)

```  

#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF BODYSHAPE TRAIT RICHNESS
#------------------------------------------------------------------------------------- #

zoom_bodyshape_pct <- zoom_species_hs_ok %>%
  group_by (loiczid)%>%
  summarize(n_tot = n(), 
  eel_pct = (sum(bodyshape==1, na.rm = TRUE)/n_tot)*100,
  elon_pct = (sum(bodyshape==2, na.rm = TRUE)/n_tot)*100,
  fusi_pct = (sum(bodyshape==3, na.rm = TRUE)/n_tot)*100,
  short_pct = (sum(bodyshape==4, na.rm = TRUE)/n_tot)*100,
  other_pct = (sum(bodyshape==5, na.rm = TRUE)/n_tot)*100,
  n_dist = n_distinct(bodyshape[!is.na(bodyshape)]))

head (zoom_bodyshape_pct)

eel_map <- subs (zoom_basemap, zoom_bodyshape_pct, by= "loiczid", which= "eel_pct")
elon_map <- subs (zoom_basemap, zoom_bodyshape_pct, by= "loiczid", which= "elon_pct")
fusi_map <- subs (zoom_basemap, zoom_bodyshape_pct, by= "loiczid", which= "fusi_pct")
short_map <- subs (zoom_basemap, zoom_bodyshape_pct, by= "loiczid", which= "short_pct")
other_map <- subs (zoom_basemap, zoom_bodyshape_pct, by= "loiczid", which= "other_pct")
dist_map <- subs (zoom_basemap, zoom_bodyshape_pct, by= "loiczid", which= "n_dist")

table (zoom_bodyshape_pct$n_dist)

breakpoints <- c(0.01,20,40,60,80,100)
colors <- c("black", "yellow","green","orange","red", "blue")

par(mfrow=c(2,3))

plot (eel_map,
      main="proportion of eel-like bodyshape spp",
      axes=FALSE, zlim=c(0,100), breaks=breakpoints,col=colors)
plot (elon_map,
      main="proportion of elongated bodyshape spp",
      axes=FALSE, zlim=c(0,100), breaks=breakpoints,col=colors)
plot (fusi_map,
      main="proportion of fusiform/normal bodyshape spp",
      axes=FALSE, zlim=c(0,100), breaks=breakpoints,col=colors)
plot (short_map,
      main="proportion of short and or deep bodyshape spp",
      axes=FALSE, zlim=c(0,100), breaks=breakpoints,col=colors)      
plot (other_map,
      main="proportion of other bodyshape spp",
      axes=FALSE, zlim=c(0,100), breaks=breakpoints,col=colors)
plot (dist_map,
      main="distinct bodyshape",
      axes=FALSE, col=rainbow(5))

#------------------------------------------------------------------------------------- #
#FUNCTIONAL DIVERSITY CALCULATIONS 
# sourcing "homemade" R functions by Sebastien Villeger (http://villeger.sebastien.free.fr/Rscripts.html)
#------------------------------------------------------------------------------------- #
#Preparing the data
# importing species raw trait values from a csv file: row names are in the column named 'am_sid'
traits_FD<-read.csv("/home/valle/github_annex/traits_FD_ok.csv", header=T)
head (traits_FD)
colnames(traits_FD)[colnames(traits_FD)=="X"] <- "am_sid"
head (traits_FD)
#we create a vector that conatins the species that are present in the selected extent
zoom_spp_list <- unique (zoom_species_hs_ok$am_sid) 
#Filter to select only the species that are present in the selected extent
zoom_traits <- filter (traits_FD, traits_FD$am_sid %in% zoom_spp_list)
unique (zoom_traits$am_sid) %>% length()#342
head (zoom_traits)
colnames (zoom_traits)
str (zoom_traits)
##Filter is keeping all levels so, we create a new dataframe to avoid that
```{r setup, include=FALSE}
zoom_traits_ok<-as.data.frame( matrix(NA, nrow(zoom_traits), ncol(zoom_traits), 
                  dimnames=list( row.names(zoom_traits), names(zoom_traits) ) ) ) 
#am_sid
zoom_traits_ok [,"am_sid"]<-as.character(zoom_traits [,"am_sid"])
zoom_traits_ok[,"am_sid"]<-factor(zoom_traits_ok[,"am_sid"], ordered = TRUE)
#class
zoom_traits_ok [,"class"]<-as.character(zoom_traits [,"class"])
zoom_traits_ok[,"class"]<-factor(zoom_traits_ok[,"class"], ordered = TRUE)
#order
zoom_traits_ok [,"order"]<-as.character(zoom_traits [,"order"])
zoom_traits_ok[,"order"]<-factor(zoom_traits_ok[,"order"], ordered = TRUE)
#family
zoom_traits_ok [,"family"]<-as.character(zoom_traits [,"family"])
zoom_traits_ok[,"family"]<-factor(zoom_traits_ok[,"family"], ordered = TRUE)
#Loo
zoom_traits_ok[,"Loo"]<-as.numeric( zoom_traits[,"Loo"] )
#K
zoom_traits_ok[,"K"]<-as.numeric( zoom_traits[,"K"] )
#Winfinity
zoom_traits_ok[,"Winfinity"]<-as.numeric( zoom_traits[,"Winfinity"] )
#tmax
zoom_traits_ok[,"tmax"]<-as.numeric( zoom_traits[,"tmax"] )
#tm
zoom_traits_ok[,"tm"]<-as.numeric( zoom_traits[,"tm"] )
#M
zoom_traits_ok[,"M"]<-as.numeric( zoom_traits[,"M"] )
#Lm
zoom_traits_ok[,"Lm"]<-as.numeric( zoom_traits[,"Lm"] )
#depthrangedeep
zoom_traits_ok[,"depthrangedeep"]<-as.numeric( zoom_traits[,"depthrangedeep"] )
#demerspelag  
zoom_traits_ok [,"demerspelag"]<-as.character(zoom_traits [,"demerspelag"])
zoom_traits_ok[,"demerspelag"]<-factor(zoom_traits_ok[,"demerspelag"], ordered = TRUE)
table (zoom_traits_ok$demerspelag)
#demersal               pelagic             reef-associated 
#            140             190              12 
#EnvTemp       
zoom_traits_ok [,"EnvTemp"]<-as.character(zoom_traits [,"EnvTemp"])
zoom_traits_ok[,"EnvTemp"]<-factor(zoom_traits_ok[,"EnvTemp"], ordered = TRUE)
table (zoom_traits_ok$EnvTemp)
#     boreal  deep-water       polar subtropical   temperate    tropical 
#           1         173           1         106          52           9 
#FeedingType   
zoom_traits_ok [,"FeedingType"]<-as.character(zoom_traits [,"FeedingType"])
zoom_traits_ok[,"FeedingType"]<-factor(zoom_traits_ok[,"FeedingType"], ordered = TRUE)
table (zoom_traits_ok$FeedingType)

#bodyshape     
zoom_traits_ok [,"bodyshape"]<-as.character(zoom_traits [,"bodyshape"])
table (zoom_traits_ok$bodyshape)
#we rename the levels into new categories names
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="eel-like"),"bodyshape"]<-"eel_like"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="elongated"),"bodyshape"]<-"elongated"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="fusiform / normal"),"bodyshape"]<-"fusiform_normal"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="short and / or deep"),"bodyshape"]<-"short_andor_deep"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="other"),"bodyshape"]<-"other"

table (zoom_traits_ok$bodyshape)

zoom_traits_ok[,"bodyshape"]<-factor(zoom_traits_ok[,"bodyshape"], levels=c("eel_like", "elongated", "fusiform_normal", "short_andor_deep", "other" ), ordered = TRUE )

#we rename into numbers
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="eel-like"),"bodyshape"]<-"1"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="elongated"),"bodyshape"]<-"2"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="fusiform / normal"),"bodyshape"]<-"3"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="short and / or deep"),"bodyshape"]<-"4"
zoom_traits_ok[which(zoom_traits_ok[,"bodyshape"]=="other"),"bodyshape"]<-"5"

zoom_traits_ok[,"bodyshape"]<-factor(zoom_traits_ok[,"bodyshape"], levels=c("1", "2", "3", "4", "5" ), ordered = TRUE )
table (zoom_traits_ok$bodyshape)

#IUCN_Code     
zoom_traits_ok [,"IUCN_Code"]<-as.character(zoom_traits [,"IUCN_Code"])
zoom_traits_ok[,"IUCN_Code"]<-factor(zoom_traits_ok[,"IUCN_Code"], ordered = TRUE)
#Resilience    
zoom_traits_ok [,"Resilience"]<-as.character(zoom_traits [,"Resilience"])
zoom_traits_ok[,"Resilience"]<-factor(zoom_traits_ok[,"Resilience"], ordered = TRUE)
#pd50          
zoom_traits_ok[,"pd50"]<-as.numeric( zoom_traits[,"pd50"] ) 
#vulnerability 
zoom_traits_ok[,"vulnerability"]<-as.numeric( zoom_traits[,"vulnerability"] ) 
#importance    
zoom_traits_ok [,"importance"]<-as.character(zoom_traits [,"importance"])
zoom_traits_ok[,"importance"]<-factor(zoom_traits_ok[,"importance"], ordered = TRUE) 
#pricecateg    
zoom_traits_ok [,"pricecateg"]<-as.character(zoom_traits [,"pricecateg"])
zoom_traits_ok[,"pricecateg"]<-factor(zoom_traits_ok[,"pricecateg"], ordered = TRUE)

str (zoom_traits_ok)
#we save the data
write.csv (zoom_traits_ok, file = "zoom_traits_eh.csv")

zoom_traits_ok<-read.csv ("zoom_traits_eh.csv", header=T, row.names = "am_sid")

#zoom_traits_ok<-read.csv("/home/valle/github_annex/zoom_traits_ok.csv", header=T, row.names = "am_sid")

head (zoom_traits_ok)
zoom_traits_ok$X <- NULL 
head (zoom_traits_ok)

#reading assemblage matrix
#zoom_assemblage <- read.csv('/home/valle/github_annex/"zoom_assemblage.csv')
zoom_assemblage <- zoom_spread_loiczid_spp
zoom_assemblage_ok <- column_to_rownames(zoom_assemblage, var = "loiczid")

#save the data
write.csv (zoom_assemblage_ok, file = "zoom_assemblage_eh.csv")

as.tibble (zoom_assemblage_ok)

# checking that species names are the same in the two matrices

sum(row.names(zoom_traits_ok) %in% colnames(zoom_assemblage_ok)) == ncol(zoom_assemblage_ok)

colnames (zoom_traits_ok)
#looking for correlations
install.packages ("corrplot")
library (corrplot)
cordb <- zoom_traits_ok %>% 
  select(-IUCN_Code, -Resilience, -vulnerability, -importance, -pricecateg, -demerspelag, -EnvTemp, -FeedingType, -class, -order, -family)
summary (cordb)
cordb_na <- drop_na(cordb) 
cor <- cor(cordb_na)
b <- corplot <- corrplot(cor,
                     type = "lower",
                     method = "ellipse",
                     diag = F,
                     outline = T,
                     tl.col = "black",
                    #tl.srt = 0,
                     cl.pos = "r",
                     cl.cex = 1,
                     cl.ratio = 0.5)

p.mat <- cor.mtest(cordb )$p

head(p.mat[, 1:9])

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
a <- corrplot(cor, method = "color", col = col(200),
         type = "upper", order = "hclust", number.cex = .7,
         addCoef.col = "black", # Add coefficient of correlation
         tl.col = "black", tl.srt = 90, # Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag = FALSE)

# species names and codes (for graphics)
species_names<-row.names(zoom_traits_ok)
species_codes<-species_names
length(unique(species_codes))==length(species_names) # OK


# setting working directory
#Desde MAZU:
my_path<-"/home/valle/github/Tutorial_FD" 
#  <= PUT direction of the folder where you saved the zipfile containing functions and data
#Desde mi portatil
#my_path<-"N:/github/Tutorial_FD"
# loading functions
setwd( paste(my_path,"/functions", sep="") ) # folder where R functions have been saved
source("quality_funct_space.R")
source("plot_funct_space.R")
source("multidimFD.R")
source("multidimFbetaD.R")

# FUNCTIONAL SPACE

setwd( paste(my_path,"/FutureSeaResults/plot_funct_space", sep="") )  # setting working directory for results

# computing all functional spaces based on dendrogram or PCoA (up to 10 axes)
# NOTE: you need to be connected to internet for a correct functioning of quality_funct_space function

#install.packages ("ape")
#install.packages("clue")
#install.packages("cluster")
#install.packages("geometry")
#install.packages("gtools")

#remove those traits that contain NAs and those traits 
summary (zoom_traits_ok)
eh_traits_noNA <- select (zoom_traits_ok, -"FeedingType", -"bodyshape", -"Resilience", -"pd50", -"depthrangedeep", -"importance", -"pricecateg", -"IUCN_Code", -"vulnerability", -"EnvTemp",-"class", -"order", -"family")
summary (eh_traits_noNA)

#we save the data
write.csv (eh_traits_noNA, file = "eh_traits_noNA.csv")

#Calculating FUNCTIONAL SPACE
qual_funct_space_eh<-quality_funct_space(eh_traits_noNA, traits_weights=NULL, nbdim=10, metric="Gower", dendro=TRUE, plot="quality_funct_space_MarineFish_eh")

qual_funct_space_eh$meanSD # => best space has 5 dimensions + have a look to the "qual_funct_space_bayofbiscay.jpeg" file in the /results folder
# mean Squared-Deviation of 0.0029 means that average deviation between Euclidean distance and Gower's distance is of (0.0024)^0.5=0.049, so it can be seen like an average error of 5%

# species coordinates in the best space
coord_fish_4D<-qual_funct_space_eh$details_funct_space$mat_coord[,1:4]

################################# HASTA AQUÍ CON LA EXTENT NUEVA
# few examples of how exploring functional space

# plot of 5D functional space => look to the jpeg file in the ".../results/plot_funct_space" folder
plot_funct_space( coord_fish_4D,col_sp="blue", pch_sp=21, nm_jpeg="Fspace_fish_4D_eh.jpeg", cex_sp=1.6, close_jpeg=TRUE   )

# looking at position of all species in the 4D space
plot_funct_space( coord_fish_5D, Faxes=c( "PC1","PC2"), Faxes_lim=c(-0.58,0.48), col_sp="blue", pch_sp=species_codes, nm_jpeg="Fspace_fish_codes_12.jpeg", cex_sp=0.6, close_jpeg=TRUE   )
plot_funct_space( coord_fish_5D, Faxes=c( "PC3","PC4"), Faxes_lim=c(-0.58,0.48), col_sp="blue", pch_sp=species_codes, nm_jpeg="Fspace_fish_codes_34.jpeg", cex_sp=0.6, close_jpeg=TRUE   )

# looking at position of 3 particular species on the 2 first PC axes
plot_funct_space( coord_fish_5D, Faxes=c( "PC1","PC2"), Faxes_lim=c(-0.55,0.45), col_sp="grey50", pch_sp="+", nm_jpeg="Fspace_fish_12_3species.jpeg", cex_sp=0.6, close_jpeg=FALSE   ) # all species plotted with crosses, not closing the jpeg
text(coord_fish_5D[c("Fis-30583","Fis-62376","Fis-23064"),1:2], c("Fis-30583","Fis-62376","Fis-23064"), cex=0.8, col="red3" ) # adding species codes for only 3 species
graphics.off() # closing the jpeg

# looking at distribution of values for one trait in the functional space
plot_funct_space( coord_fish_5D, Faxes=c( "PC1","PC2"), Faxes_lim=c(-0.6,0.6), col_sp=as.factor(zoom_traits_ok_NA[,"demerspelag"]), pch_sp=21, nm_jpeg="Fspace_fish_12_demerspelag.jpeg", cex_sp=0.6, close_jpeg=FALSE   ) # points color defined according to values of this categorical trait, not closing the jpeg
plot(0:7,0:7,type="n", axes=F,xlab="", ylab="") # empty plot for legend on the right panel
points(rep(1.8,7), 1:7, pch=21, cex=1, bg=as.factor(levels(zoom_traits_ok_NA[,"demerspelag"])) ) 
text(rep(2,7), 1:7, levels(zoom_traits_ok_NA[,"demerspelag"]), cex=1, adj=0 ) 
graphics.off() # closing the jpeg

# looking at distribution of values for one trait in the functional space
plot_funct_space( coord_fish_5D, Faxes=c( "PC1","PC2"), Faxes_lim=c(-0.6,0.6), col_sp=as.factor(zoom_traits_ok_NA[,"EnvTemp"]), pch_sp=21, nm_jpeg="Fspace_fish_12_EnvTemp.jpeg", cex_sp=0.6, close_jpeg=FALSE   ) # points color defined according to values of this categorical trait, not closing the jpeg
plot(0:6,0:6,type="n", axes=F,xlab="", ylab="") # empty plot for legend on the right panel
points(rep(1.8,6), 1:6, pch=21, cex=1, bg=as.factor(levels(zoom_traits_ok_NA[,"EnvTemp"])) ) 
text(rep(2,6), 1:6, levels(zoom_traits_ok_NA[,"EnvTemp"]), cex=1, adj=0 ) 
graphics.off() # closing the jpeg

# looking at distribution of values for one trait in the functional space
plot_funct_space( coord_fish_5D, Faxes=c( "PC1","PC2"), Faxes_lim=c(-0.6,0.6), col_sp=as.factor(zoom_traits_ok_NA[,"class"]), pch_sp=21, nm_jpeg="Fspace_fish_12_class.jpeg", cex_sp=0.6, close_jpeg=FALSE   ) # points color defined according to values of this categorical trait, not closing the jpeg
plot(0:6,0:6,type="n", axes=F,xlab="", ylab="") # empty plot for legend on the right panel
points(rep(1.8,4), 1:4, pch=21, cex=1, bg=as.factor(levels(zoom_traits_ok_NA[,"class"])) ) 
text(rep(2,4), 1:4, levels(zoom_traits_ok_NA[,"class"]), cex=1, adj=0 ) 
graphics.off() # closing the jpeg

##############################################################################################
# illustration of bias induced by functional dendrogram or space of low dimensionality

# functional differences between "Fis-30583","Fis-62376","Fis-23064"

# trait values
zoom_traits_ok_NA[c("Fis-30583","Fis-62376","Fis-23064"),]

# raw distance based on traits values 
round(as.matrix(qual_funct_space$details_funct_space$mat_dissim)[c("Fis-30583","Fis-62376","Fis-23064"),c("Fis-30583","Fis-62376","Fis-23064")],3)
# => Fis-30583 is almost >2 times closer to Fis-23064 than to Fis-62376

##################################
# plot of best functional dendrogram
plot(qual_funct_space_eh$details_funct_space$best_tree, sub="", main="UPGMA", xlab="", h=-1)
# look at position of Fis-30583 on the tree

# cophenetic distance on best dendrogram 
round(as.matrix(qual_funct_space$details_funct_space$dist_raw$t_UPGMA)[c("Fis-30583","Fis-62376","Fis-23064"),c("Fis-30583","Fis-62376","Fis-23064")],3) 
# Fis-30583 as far to Fis-23064 than to Fis-62376 => dendrogram overestimates distance between some pairs of species actually close

# most extreme species is the Fis-62376
round(as.matrix(qual_funct_space$details_funct_space$dist_raw$t_UPGMA)["Fis-62376",],2) 


round( as.matrix(qual_funct_space$details_funct_space$mat_dissim)["Fis-62376",],2)

# correlation between distance on the dendrogram vs Gower distance for pairs between Fis-62376 and other fish
plot( x=as.matrix(qual_funct_space$details_funct_space$mat_dissim)["Fis-62376",], y=as.matrix(qual_funct_space$details_funct_space$dist_raw$t_UPGMA)["Fis-62376",], xlab="Gower Dist on trait values", ylab="Dist on UPGMA dendro", pch=16, cex=0.7  )
# no strong correlation : 2 horizontal lines of points with the points close to y=0.4 corresponding to pineapple vs ( banana, litchi or mango)
# and the points at y=0.56 corresponding to distance between pineapple and all other fish
# => dendrogram overestimates some distances and homogenize them (lower variability than with Gower distance)

######hasta aquí


##################################
# distance in the 2D space 
round(as.matrix(qual_funct_space$details_funct_space$dist_raw$m_2D)[c("cherry","lime","lemon"),c("cherry","lime","lemon")],3)
# lime 2.4 times closer to lemon than to cherry (instead of 3 based on trait values)
# => look at position of these 3 species on PC1-PC2 on "Fspace_fish_codes_12.jpeg", cherry is not that far from both Citrus species

##################################
# distance in the 5D space 
round(as.matrix(qual_funct_space$details_funct_space$dist_raw$m_5D)[c("cherry","lime","lemon"),c("cherry","lime","lemon")],3)
# lime 2.4 times closer to lemon than to cherry
# => look at position of these 3 species on PC3-PC4 on "Fspace_fish_5D.jpeg", cherry is far from both Citrus species

# correlation between distance in the 5D space vs Gower distance for pairs between pineapple and other fish
plot( x=as.matrix(qual_funct_space$details_funct_space$mat_dissim)["pineapple",], y=as.matrix(qual_funct_space$details_funct_space$dist_raw$m_5D)["pineapple",], xlab="Gower Dist on trait values", ylab="Dist in 5D space", pch=16, cex=0.7  )
# strong correlation between Euclidean distance in the 5D space and Gower distance => no overestimation, no homogenization

##############################################################################################
##############################################################################################

# MULTIDIMENSIONAL FUNCTIONAL DIVERISTY INDICES

# computing Functional diversity indices with plots of FD indices put in a subfolder named plot_FD
as.tibble (zoom_assemblage_ok)

matrix_zoom_assemblage_ok <- data.matrix(zoom_assemblage_ok)

FD_assemblages_eh2<-multidimFD(coord_fish_4D, matrix_zoom_assemblage_ok , check_species_pool=TRUE, verb=TRUE,
                      nm_asb_plot=row.names(matrix_zoom_assemblage_ok), 
                      folder_plot= paste(my_path,"/results/plot_FD/eh", sep=""),
                      Faxes_plot=colnames(coord_fish_4D)[1:4], Faxes_nm_plot=colnames(coord_fish_4D)[1:4],
                      plot_pool=TRUE, col_bg="grey90", col_sp_pool="grey30",  pch_sp_pool="+", cex_sp_pool=1, 
                      pch_sp=21, col_sp="#1E90FF", transp=50 )

# printing results = rounded FD indices values
round(FD_assemblages_eh,3)

#plotting results

rich_fd <- subset (FD_assemblages_eh, select=c("Nb_sp", "FRic", "FDiv", "FEve", "FDis", "FSpe", "FOri"))
rich_fd <- as.data.frame(rich_fd )

#FRic=Functional richness = convex hull volume

ggplot(rich_fd, aes(x =Nb_sp, y = FRic )) +
  geom_point()

#FDiv=Functional Divergence

ggplot(rich_fd, aes(x =Nb_sp, y = FDiv)) +
  geom_point()

#FEve=Functional Evenness

ggplot(rich_fd, aes(x =Nb_sp, y = FEve)) +
  geom_point()

#FDis=Functional Dispersion: abundance-weighted mean distance to abundance-weighted centroid scaled by maximum value possible given species pool (i.e. the two most distant species have half of total weight)

ggplot(rich_fd, aes(x =Nb_sp, y = FDis, na.rm = TRUE)) +
  geom_point()

#FSpe=Functional Specialization: abundance-weighted mean distance to centroid of the global pool of species scaled by maximum value possible given species pool (i.e. an assmeblage hosting only the most specialized species)

ggplot(rich_fd, aes(x =Nb_sp, y = FSpe)) +
  geom_point()

#FOri=Functional Originality : abundance-weighted mean distance to nearest neighbour in the global pool of species scaled by maximum value possible given species pool (i.e. an assmeblage hosting only the most original species)
ggplot(rich_fd, aes(x =Nb_sp, y = FOri)) +
  geom_point()

#categories for congruence calculations see Martín-Regalado et al 2019

summary (rich_fd$Nb_sp)

#selecting the category ranges manually: 
#rich_fd$categoryNb_sp[rich_fd$Nb_sp < 500] <- "low"
#rich_fd$categoryNb_sp[rich_fd$Nb_sp > 501 & rich_fd$Nb_sp < 1000] <- "middle"
#rich_fd$categoryNb_sp[rich_fd$Nb_sp > 1001] <- "high"

#selecting the category ranges by quantile:
xs<-quantile(rich_fd$Nb_sp,c(0,1/3,2/3,1))
xs[1]<- xs[1]-.00005
df1 <- rich_fd %>% mutate(category=cut(Nb_sp, breaks=xs, labels=c("low","middle","high")))
boxplot(df1$Nb_sp~df1$category,col=3:5)

#selecting the category ranges by quantile:
xsfd<-quantile(rich_fd$FRic,c(0,1/3,2/3,1),na.rm=TRUE)
xsfd[1]<- xsfd[1]-.00005
df2 <- df1 %>% mutate(category2=cut(FRic, breaks=xsfd, labels=c("low","middle","high")))
boxplot(df2$FRic~df2$category2,col=3:5)

#CALCULATING CONGRUENCE LEVEL

#high if identical categories
high if Nb_sp == high & FRic == high or Nb_sp == middle & FRic == middle or Nb_sp == low & FRic == low

#moderate if contiguous
moderate if Nb_sp == high & FRic == middle or Nb_sp == middle & FRic == high or Nb_sp == moderate & FRic == low or Nb_sp == low & FRic == moderate

#low if extreme
low if Nb_sp == high & FRic == low or Nb_sp == low & FRic == high 


#FUNCIONAL DIvergency
loiczid <- rownames(rich_fd)
rich_fd <- cbind(loiczid=loiczid, rich_fd)

zoom_FD_map <- subs (zoom_basemap, rich_fd, by= "loiczid", which= "FDiv")

plot (zoom_FD_map, zlim=c(0.8,1), col=rainbow(5))

#FUNCIONAL richness MAP

zoom_FDric_map <- subs (zoom_basemap, rich_fd, by= "loiczid", which= "FRic")

plot (zoom_FDric_map, zlim=c(0,1))


zoom_spp2_map <- subs (zoom_basemap, rich_fd, by= "loiczid", which= "Nb_sp")
plot (zoom_spp2_map)

# look to the folder "../results/plot_FD"
##############################################################################################
# MULTIDIMENSIONAL FUNCTIONAL BETA DIVERISTY INDICES

# occurences from species weights
occ_fish_assemblages<-matrix_zoom_assemblage_ok_100
occ_fish_assemblages[which(occ_fish_assemblages>0)]<-1

# indices computation for all pairs of asssemblages, plot only on 3 first axes
install.packages ("betapart")
FbetaD_assemblages<-multidimFbetaD ( coord=coord_fish_5D,  occ=occ_fish_assemblages,  check_species_pool=FALSE, verb=TRUE,
                  nm_asb_plot=row.names(occ_fish_assemblages) , folder_plot= paste(my_path,"/results/plot_FbetaD", sep="") , 
                  Faxes_plot=colnames(coord_fish_5D)[1:3] , Faxes_nm_plot=colnames(coord_fish_5D)[1:3] )

# printing results
FbetaD_assemblages

# look to the folder "../results/plot_FbetaD"

##############################################################################################
# saving R objects

setwd( paste(my_path,"/results", sep="") ) # setting working directory for results

save(qual_funct_space_clear, file="qual_funct_space_clear")


save(zoom_traits_ok_NA, file="zoom_traits_ok_NA")
save(species_codes, file="species_codes")

save(coord_fish_5D, file="coord_fish_5D")
save(FD_assemblages, file="FD_assemblages")

save(weight_fish_baskets, file="weight_fish_baskets")
save(FbetaD_baskets, file="FbetaD_baskets")

##############################################################################################
# Going one step further: example of how running a null model
# here we want to test H0: Basket_4 results from a random sorting of species in the regional species pool (i.e. all species) given its species richness (SR). To keep the example simple, will focus only on FRic index so taking into account only species composition, not species weights.


# picking assemblages of SR(basket_4) species at random among the 25 fish
nbrep<-99 # number of replicates
SR_basket_4<-FD_baskets["basket_4","Nb_sp"]
SR_basket_4 # 8 species`

# empty matrix to store simulated species occurences
basket_4_H0<-matrix(0, nbrep, ncol(weight_fish_baskets), dimnames=list(1:nbrep, colnames(weight_fish_baskets) ) )
for (k in 1:nbrep)
{
  basket_4_H0[k, sample(  colnames(weight_fish_baskets), SR_basket_4) ]<-1 # random sorting of species
}# end of k

# computing FD indices on these assemblages, check_species_pool=FALSE since by chance some species could be never picked but this is not an issue
FD_basket_4_H0<-multidimFD(coord_fish_5D, basket_4_H0,  check_species_pool=FALSE )

# comparing observed and expected values under H0 using SES and p-value metrics
SES_FRic_basket_4<- (FD_baskets["basket_4","FRic"]-mean(FD_basket_4_H0[,"FRic"]) ) / sd(FD_basket_4_H0[,"FRic"])
SES_FRic_basket_4 # SES<(-1) means that observed FRic is lower than expected
  
pvalue_FRic_basket_4<- length(which(FD_baskets["basket_4","FRic"]<=FD_basket_4_H0[,"FRic"]))/ ( length(FD_basket_4_H0[,"FRic"]) +1 )
pvalue_FRic_basket_4 # p-value >0.975 => FRic is significantly lower than expected under H0


##############################################################################################
# END
##############################################################################################
