---
title: "1_preparing homemade dataset"
author: "Mireia Valle"
date: "12/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#BUILDING A DATA FRAME WITH LOICZID VALUE EQUAL TO THE EXTENT WE ARE INTERESTED IN 
##We first create the global map
```{r, include = FALSE}

basemap <- raster (ext= extent (c(-180,180, -90, 90)), res=0.5)
values (basemap) <- 1:length (basemap)
basemap
plot (basemap)
```
##We then create a data frame with long, lat and loiczid columns
```{r, include = FALSE}

base_df <- rasterToPoints(basemap) %>%
  as.data.frame() %>%
  setNames(c('long', 'lat', 'loiczid')) %>%
  mutate(loiczid = as.integer(loiczid))

head (base_df)
```
##We filter the data frame to the extent we are interested in: (ext= extent (c(-6,4, 36, 44))
```{r setup}
zoom_base_df1 <- filter(base_df, between(base_df$long, -10, 4))
zoom_base_df <- filter(zoom_base_df1, between(zoom_base_df1$lat, 36, 44))
```
##We remove the first filtered dataframe
```{r setup}
rm(zoom_base_df1)
```
##We look at our new zoomed data frame
```{r setup}
head (zoom_base_df)
str (zoom_base_df)
unique (zoom_base_df$loiczid) %>% length ()#448
summary (zoom_base_df$lat)
summary (zoom_base_df$long)
```
##We create a vector with loiczid values for the extent we are interested in
```{r setup}
zoom_loiczid<- zoom_base_df$loiczid
```
##We filter "species_hs" dataframe in order to get whole information about the different species that are located in the are we have selected
```{r setup}
#species_hs <- run 7_Mapping traits.Rmd from line 38 to line 76
head (species_hs)
```
##We filter the spcies_hs dataframe by the loiczid that we want to analyse
```{r setup}
zoom_species_hs <- filter (species_hs, species_hs$loiczid %in% zoom_loiczid)
unique (zoom_species_hs$loiczid) %>% length() #209
unique (zoom_species_hs$am_sid) %>% length()#1184
```

##We filter the probability value to 1
```{r setup}
occurrence <- filter(zoom_species_hs, prob ==1)
unique (occurrence$am_sid) %>% length()#844

head (occurrence)
colnames (occurrence)
str (occurrence)

zoom_species_hs <- occurrence
rm (occurrence)
```

##Filter is keeping all levels so, we create a new dataframe to avoid that
```{r setup, include=FALSE}
zoom_species_hs_ok<-as.data.frame( matrix(NA, nrow(zoom_species_hs), ncol(zoom_species_hs), 
                  dimnames=list( row.names(zoom_species_hs), names(zoom_species_hs) ) ) ) 
#am_sid
zoom_species_hs_ok [,"am_sid"]<-as.character(zoom_species_hs [,"am_sid"])
zoom_species_hs_ok[,"am_sid"]<-factor(zoom_species_hs_ok[,"am_sid"], ordered = TRUE)
#loiczid
zoom_species_hs_ok [,"loiczid"]<-as.integer(zoom_species_hs [,"loiczid"])
#prob
zoom_species_hs_ok[,"prob"]<-as.numeric(zoom_species_hs[,"prob"] )
#class
zoom_species_hs_ok [,"class"]<-as.character(zoom_species_hs [,"class"])
zoom_species_hs_ok[,"class"]<-factor(zoom_species_hs_ok[,"class"], ordered = TRUE)
#order
zoom_species_hs_ok [,"order"]<-as.character(zoom_species_hs [,"order"])
zoom_species_hs_ok[,"order"]<-factor(zoom_species_hs_ok[,"order"], ordered = TRUE)
#family
zoom_species_hs_ok [,"family"]<-as.character(zoom_species_hs [,"family"])
zoom_species_hs_ok[,"family"]<-factor(zoom_species_hs_ok[,"family"], ordered = TRUE)
#Loo
zoom_species_hs_ok[,"Loo"]<-as.numeric( zoom_species_hs[,"Loo"] )
#K
zoom_species_hs_ok[,"K"]<-as.numeric( zoom_species_hs[,"K"] )
#Winfinity
zoom_species_hs_ok[,"Winfinity"]<-as.numeric( zoom_species_hs[,"Winfinity"] )
#tmax
zoom_species_hs_ok[,"tmax"]<-as.numeric( zoom_species_hs[,"tmax"] )
#tm
zoom_species_hs_ok[,"tm"]<-as.numeric( zoom_species_hs[,"tm"] )
#M
zoom_species_hs_ok[,"M"]<-as.numeric( zoom_species_hs[,"M"] )
#Lm
zoom_species_hs_ok[,"Lm"]<-as.numeric( zoom_species_hs[,"Lm"] )
#depthrangedeep
zoom_species_hs_ok[,"depthrangedeep"]<-as.numeric( zoom_species_hs[,"depthrangedeep"] )
#demerspelag  
zoom_species_hs_ok [,"demerspelag"]<-as.character(zoom_species_hs [,"demerspelag"])
zoom_species_hs_ok[,"demerspelag"]<-factor(zoom_species_hs_ok[,"demerspelag"], ordered = TRUE)
#EnvTemp       
zoom_species_hs_ok [,"EnvTemp"]<-as.character(zoom_species_hs [,"EnvTemp"])
zoom_species_hs_ok[,"EnvTemp"]<-factor(zoom_species_hs_ok[,"EnvTemp"], ordered = TRUE)
#FeedingType   
zoom_species_hs_ok [,"FeedingType"]<-as.character(zoom_species_hs [,"FeedingType"])
zoom_species_hs_ok[,"FeedingType"]<-factor(zoom_species_hs_ok[,"FeedingType"], ordered = TRUE)
#bodyshape     
zoom_species_hs_ok [,"bodyshape"]<-as.character(zoom_species_hs [,"bodyshape"])
#we rename the levels into new categories names
table (zoom_species_hs_ok$bodyshape)

zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="eel-like"),"bodyshape"]<-"eel_like"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="elongated"),"bodyshape"]<-"elongated"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="fusiform / normal"),"bodyshape"]<-"fusiform_normal"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="short and / or deep"),"bodyshape"]<-"short_andor_deep"
zoom_species_hs_ok[which(zoom_species_hs_ok[,"bodyshape"]=="other"),"bodyshape"]<-"other"

zoom_species_hs_ok[,"bodyshape"]<-factor(zoom_species_hs_ok[,"bodyshape"], levels=c("eel_like", "elongated", "fusiform_normal", "short_andor_deep", "other" ), ordered = TRUE )

table (zoom_species_hs_ok$bodyshape)

#IUCN_Code     
zoom_species_hs_ok [,"IUCN_Code"]<-as.character(zoom_species_hs [,"IUCN_Code"])
zoom_species_hs_ok[,"IUCN_Code"]<-factor(zoom_species_hs_ok[,"IUCN_Code"], ordered = TRUE)
#Resilience    
zoom_species_hs_ok [,"Resilience"]<-as.character(zoom_species_hs [,"Resilience"])
zoom_species_hs_ok[,"Resilience"]<-factor(zoom_species_hs_ok[,"Resilience"], ordered = TRUE)
#pd50          
zoom_species_hs_ok[,"pd50"]<-as.numeric( zoom_species_hs[,"pd50"] ) 
#vulnerability 
zoom_species_hs_ok[,"vulnerability"]<-as.numeric( zoom_species_hs[,"vulnerability"] ) 
#importance    
zoom_species_hs_ok [,"importance"]<-as.character(zoom_species_hs [,"importance"])
zoom_species_hs_ok[,"importance"]<-factor(zoom_species_hs_ok[,"importance"], ordered = TRUE) 
#pricecateg    
zoom_species_hs_ok [,"pricecateg"]<-as.character(zoom_species_hs [,"pricecateg"])
zoom_species_hs_ok[,"pricecateg"]<-factor(zoom_species_hs_ok[,"pricecateg"], ordered = TRUE)

str (zoom_species_hs_ok)

```
##We remove the original data filtered dataset
```{r setup}
rm (zoom_species_hs)
```
##We create a subset of the zoom_species_hs_ok in order to create the assemblage matrix (loiczid/spps probability)
```{r setup}
zoom_species_hs_subset <- subset(zoom_species_hs_ok , select=c("am_sid", "loiczid","prob"))

head (zoom_species_hs_subset)
as.tibble (zoom_species_hs_subset)

unique (zoom_species_hs_subset$am_sid) %>% length()#691

#we spread the data
zoom_spread_loiczid_spp <- zoom_species_hs_subset %>%
    spread(am_sid, prob, fill = 0)

as.tibble (zoom_spread_loiczid_spp)
#we save the data
write.csv (zoom_spread_loiczid_spp, file = "zoom_assemblage_iberian_peninsula.csv")
```
#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF ALL SPECIES IN THE SELECTED EXTEND ()
#------------------------------------------------------------------------------------- #
```{r setup}
zoom_spp <-  zoom_species_hs_ok %>%
  group_by (loiczid)%>%
  summarize(n_spp = n())
  
head (zoom_spp)
unique(zoom_species_hs_ok$am_sid) %>% length()#844
unique(zoom_species_hs_ok$loiczid) %>% length()#209
unique(zoom_spp$loiczid) %>% length()#209
```
##We create the zoom map
```{r setup}
zoom_basemap <- raster (ext= extent (c(-10,4, 36, 44)), res=0.5)
values (zoom_basemap) <- zoom_loiczid
zoom_basemap
plot (zoom_basemap)

zoom_spp_map <- subs (zoom_basemap, zoom_spp, by= "loiczid", which= "n_spp")

plot (zoom_spp_map)
```

##--------------------- SMALLER EXTENT: CANTABRIAN COAST (cc) -----------------------------------------------------##

##We filter the data frame to the extent we are interested in: Cantabrian Coast: (ext= extent (c(-8.5,4, 43, 43.5))
```{r setup}
zoom_base_df_cc1 <- filter(base_df, between(base_df$long, -8.5,4))
zoom_base_df_cc <- filter(zoom_base_df_cc1, between(zoom_base_df_cc1$lat, 43, 43.5))
```
##We remove the first filtered dataframe
```{r setup}
rm(zoom_base_df_cc1)
```
##We look at our new zoomed data frame
```{r setup}
head (zoom_base_df_cc)
str (zoom_base_df_cc)
unique (zoom_base_df_cc$loiczid) %>% length ()
summary (zoom_base_df_cc$lat)
summary (zoom_base_df_cc$long)
```
##We create a vector with loiczid values for the extent we are interested in
```{r setup}
zoom_loiczid_cc<- zoom_base_df_cc$loiczid
head (zoom_loiczid_cc)
```

##We filter "species_hs" dataframe in order to get whole information about the different species that are located in the are we have selected
```{r setup}
#species_hs <- run 7_Mapping traits.Rmd from line 38 to line 76
head (species_hs)
```
##We filter the spcies_hs dataframe by the loiczid that we want to analyse
```{r setup}
zoom_species_hs_cc <- filter (species_hs, species_hs$loiczid %in% zoom_loiczid_cc)
unique (zoom_species_hs_cc$loiczid) %>% length() #10
unique (zoom_species_hs_cc$am_sid) %>% length()#724
```

##We filter the probability value to 1
```{r setup}
occurrence <- filter(zoom_species_hs_cc, prob ==1)
unique (occurrence$am_sid) %>% length()#279

head (occurrence)
colnames (occurrence)
str (occurrence)

zoom_species_hs_cc <- occurrence
rm (occurrence)
```

##Filter is keeping all levels so, we create a new dataframe to avoid that
```{r setup, include=FALSE}
zoom_species_hs_ok_cc<-as.data.frame( matrix(NA, nrow(zoom_species_hs_cc), ncol(zoom_species_hs_cc), 
                  dimnames=list( row.names(zoom_species_hs_cc), names(zoom_species_hs_cc) ) ) ) 
#am_sid
zoom_species_hs_ok_cc [,"am_sid"]<-as.character(zoom_species_hs_cc [,"am_sid"])
zoom_species_hs_ok_cc[,"am_sid"]<-factor(zoom_species_hs_ok_cc[,"am_sid"], ordered = TRUE)
#loiczid
zoom_species_hs_ok_cc [,"loiczid"]<-as.integer(zoom_species_hs_cc [,"loiczid"])
#prob
zoom_species_hs_ok_cc[,"prob"]<-as.numeric(zoom_species_hs_cc[,"prob"] )
#class
zoom_species_hs_ok_cc [,"class"]<-as.character(zoom_species_hs_cc [,"class"])
zoom_species_hs_ok_cc[,"class"]<-factor(zoom_species_hs_ok_cc[,"class"], ordered = TRUE)
#order
zoom_species_hs_ok_cc [,"order"]<-as.character(zoom_species_hs_cc [,"order"])
zoom_species_hs_ok_cc[,"order"]<-factor(zoom_species_hs_ok_cc[,"order"], ordered = TRUE)
#family
zoom_species_hs_ok_cc [,"family"]<-as.character(zoom_species_hs_cc [,"family"])
zoom_species_hs_ok_cc[,"family"]<-factor(zoom_species_hs_ok_cc[,"family"], ordered = TRUE)
#Loo
zoom_species_hs_ok_cc[,"Loo"]<-as.numeric( zoom_species_hs_cc[,"Loo"] )
#K
zoom_species_hs_ok_cc[,"K"]<-as.numeric( zoom_species_hs_cc[,"K"] )
#Winfinity
zoom_species_hs_ok_cc[,"Winfinity"]<-as.numeric( zoom_species_hs_cc[,"Winfinity"] )
#tmax
zoom_species_hs_ok_cc[,"tmax"]<-as.numeric( zoom_species_hs_cc[,"tmax"] )
#tm
zoom_species_hs_ok_cc[,"tm"]<-as.numeric( zoom_species_hs_cc[,"tm"] )
#M
zoom_species_hs_ok_cc[,"M"]<-as.numeric( zoom_species_hs_cc[,"M"] )
#Lm
zoom_species_hs_ok_cc[,"Lm"]<-as.numeric( zoom_species_hs_cc[,"Lm"] )
#depthrangedeep
zoom_species_hs_ok_cc[,"depthrangedeep"]<-as.numeric( zoom_species_hs_cc[,"depthrangedeep"] )
#demerspelag  
zoom_species_hs_ok_cc [,"demerspelag"]<-as.character(zoom_species_hs_cc [,"demerspelag"])
zoom_species_hs_ok_cc[,"demerspelag"]<-factor(zoom_species_hs_ok_cc[,"demerspelag"], ordered = TRUE)
#EnvTemp       
zoom_species_hs_ok_cc [,"EnvTemp"]<-as.character(zoom_species_hs_cc [,"EnvTemp"])
zoom_species_hs_ok_cc[,"EnvTemp"]<-factor(zoom_species_hs_ok_cc[,"EnvTemp"], ordered = TRUE)
#FeedingType   
zoom_species_hs_ok_cc [,"FeedingType"]<-as.character(zoom_species_hs_cc [,"FeedingType"])
zoom_species_hs_ok_cc[,"FeedingType"]<-factor(zoom_species_hs_ok_cc[,"FeedingType"], ordered = TRUE)
#bodyshape     
zoom_species_hs_ok_cc [,"bodyshape"]<-as.character(zoom_species_hs_cc [,"bodyshape"])
#we rename the levels into new categories names
table (zoom_species_hs_ok_cc$bodyshape)

zoom_species_hs_ok_cc[which(zoom_species_hs_ok_cc[,"bodyshape"]=="eel-like"),"bodyshape"]<-"eel_like"
zoom_species_hs_ok_cc[which(zoom_species_hs_ok_cc[,"bodyshape"]=="elongated"),"bodyshape"]<-"elongated"
zoom_species_hs_ok_cc[which(zoom_species_hs_ok_cc[,"bodyshape"]=="fusiform / normal"),"bodyshape"]<-"fusiform_normal"
zoom_species_hs_ok_cc[which(zoom_species_hs_ok_cc[,"bodyshape"]=="short and / or deep"),"bodyshape"]<-"short_andor_deep"
zoom_species_hs_ok_cc[which(zoom_species_hs_ok_cc[,"bodyshape"]=="other"),"bodyshape"]<-"other"

zoom_species_hs_ok_cc[,"bodyshape"]<-factor(zoom_species_hs_ok_cc[,"bodyshape"], levels=c("eel_like", "elongated", "fusiform_normal", "short_andor_deep", "other" ), ordered = TRUE )

table (zoom_species_hs_ok_cc$bodyshape)

#IUCN_Code     
zoom_species_hs_ok_cc [,"IUCN_Code"]<-as.character(zoom_species_hs_cc [,"IUCN_Code"])
zoom_species_hs_ok_cc[,"IUCN_Code"]<-factor(zoom_species_hs_ok_cc[,"IUCN_Code"], ordered = TRUE)

#Resilience    
zoom_species_hs_ok_cc [,"Resilience"]<-as.character(zoom_species_hs_cc [,"Resilience"])
zoom_species_hs_ok_cc[,"Resilience"]<-factor(zoom_species_hs_ok_cc[,"Resilience"], ordered = TRUE)
#pd50          
zoom_species_hs_ok_cc[,"pd50"]<-as.numeric( zoom_species_hs_cc[,"pd50"] ) 
#vulnerability 
zoom_species_hs_ok_cc[,"vulnerability"]<-as.numeric( zoom_species_hs_cc[,"vulnerability"] ) 
#importance    
zoom_species_hs_ok_cc [,"importance"]<-as.character(zoom_species_hs_cc [,"importance"])
zoom_species_hs_ok_cc[,"importance"]<-factor(zoom_species_hs_ok_cc[,"importance"], ordered = TRUE) 
#pricecateg    
zoom_species_hs_ok_cc [,"pricecateg"]<-as.character(zoom_species_hs_cc [,"pricecateg"])
zoom_species_hs_ok_cc[,"pricecateg"]<-factor(zoom_species_hs_ok_cc[,"pricecateg"], ordered = TRUE)

str (zoom_species_hs_ok_cc)

```
##We remove the original data filtered dataset
```{r setup}
rm (zoom_species_hs_cc)
```
##We create a subset of the zoom_species_hs_ok_cc in order to create the assemblage matrix (loiczid/spps probability)
```{r setup}
zoom_species_hs_cc_subset <- subset(zoom_species_hs_ok_cc , select=c("am_sid", "loiczid","prob"))

head (zoom_species_hs_cc_subset)
as.tibble (zoom_species_hs_cc_subset)

unique (zoom_species_hs_cc_subset$am_sid) %>% length()#279

#we spread the data
zoom_spread_loiczid_spp_cc <- zoom_species_hs_cc_subset %>%
    spread(am_sid, prob, fill = 0)

as.tibble (zoom_spread_loiczid_spp_cc)
#we save the data
write.csv (zoom_spread_loiczid_spp, file = "zoom_assemblage_cantabrian_coast.csv")
```
#------------------------------------------------------------------------------------- #
#MAPPING THE DISTRIBUTION OF ALL SPECIES IN THE SELECTED EXTEND ()
#------------------------------------------------------------------------------------- #
```{r setup}
zoom_spp_cc <-  zoom_species_hs_ok_cc %>%
  group_by (loiczid)%>%
  summarize(n_spp = n())

head(zoom_spp_cc)
  
head (zoom_spp_cc)
unique(zoom_species_hs_ok_cc$am_sid) %>% length()#279
unique(zoom_species_hs_ok_cc$loiczid) %>% length()#10
unique(zoom_spp_cc$loiczid) %>% length()#10
```
##We create the zoom map
```{r setup}
zoom_basemap_cc <- raster (ext= extent (c(-8.5,4, 43, 43.5)), res=0.5)
values (zoom_basemap_cc) <- zoom_loiczid_cc
zoom_basemap_cc
plot (zoom_basemap_cc)

zoom_spp_map_cc <- subs (zoom_basemap_cc, zoom_spp_cc, by= "loiczid", which= "n_spp")

plot (zoom_spp_map_cc)
```
#---------------SUBSETTING NUMBER OF SPECIES ------------------------------------------------------------------------#
#We want to create a database containing 50 random species-------#
#we read the traits dataframe
```{r}
traits_FD<- read.csv ("/home/valle/github_annex/traits_FD_ok.csv") 
as.tibble (traits_FD)
colnames (traits_FD)
summary (traits_FD)
colnames(traits_FD)[colnames(traits_FD)=="X"] <- "am_sid"
as.tibble (traits_FD)
```
#we create a vector with those species that appear at the cantabrian coast (cc)
```{r}
am_sid_cc <- unique(zoom_species_hs_ok_cc$am_sid)
am_sid_cc
```
#we filter the traits dataframe to select only the species that are present in the cc
```{r}
traits_sp_cc <- filter (traits_FD, traits_FD$am_sid %in% am_sid_cc)
head (traits_sp_cc)
unique(traits_sp_cc$am_sid) %>% length()#279
```
#now we select 50 random rows in order to create a easy to handle dataframe
```{r}
traits_sp_cc_50_random <-traits_sp_cc[sample(nrow(traits_sp_cc),50),]
```
#we save the data
```{r}
write.csv (traits_sp_cc_50_random, file = "traits_matrix_cc_50_random.csv")
```
#and we select the same 50 species from the hs dataframe
```{r}
head (zoom_species_hs_cc_subset)
as.tibble (zoom_species_hs_cc_subset)

head (traits_sp_cc_50_random)

species_hs_cc_50_random <- filter (zoom_species_hs_cc_subset, zoom_species_hs_cc_subset$am_sid %in% traits_sp_cc_50_random$am_sid)

head (species_hs_cc_50_random)

unique (species_hs_cc_50_random$am_sid ) %>%  length ()
```
#we spread the data
```{r}
species_hs_cc_50_random_spread <- species_hs_cc_50_random %>%
    spread(am_sid, prob, fill = 0)

as.tibble (species_hs_cc_50_random_spread)

unique (species_hs_cc_50_random_spread$loiczid) %>%  length ()
```
#we save the data
```{r}

write.csv (species_hs_cc_50_random_spread, file = "presence_absence_matrix_cc_50_random.csv")
```








